\subsubsection{{\sf assign}: Row variant}

Assign the contents a vector to a subset of elements in one row of a matrix. 
Note that since the output cannot be transposed, a different variant of
{\sf assign} is provided to assign to a column of a matrix.

\paragraph{\syntax}

\begin{verbatim}
        GrB_Info GrB_assign(GrB_Matrix            C,
                            const GrB_Vector      mask,
                            const GrB_BinaryOp    accum,
                            const GrB_Vector      u,
                            GrB_Index             row_index,
                            const GrB_Index      *col_indices,
                            const GrB_Index       ncols,
                            const GrB_Descriptor  desc);
\end{verbatim}

\paragraph{Parameters}

\begin{itemize}[leftmargin=1.1in]
    \item[{\sf C}]    ({\sf INOUT}) An existing GraphBLAS Matrix.  On input,
    the matrix provides values that may be accumulated with the result of the
    assign operation.  On output, this matrix holds the results of the
    operation.

    \item[{\sf mask}] ({\sf IN}) An optional ``write'' mask that controls which
    results from this operation are stored into the specified row of the output matrix {\sf C}. The 
    mask dimensions must match those of a single row of the matrix {\sf C} and the domain of the 
    {\sf Mask} matrix must be of type {\sf bool} or any of the predefined 
    ``built-in'' types in Table~\ref{Tab:PredefinedTypes}.  If the default
    vector is desired (\ie, with correct dimensions filled with {\sf true}), 
    {\sf GrB\_NULL} should be specified.

    \item[{\sf accum}] ({\sf IN}) An optional binary operator used for accumulating
    entries into existing {\sf C} entries. If assignment rather than accumulation is
    desired, {\sf GrB\_NULL} should be specified.

    \item[{\sf u}]       ({\sf IN}) The GraphBLAS vector whose contents are 
    assigned to (a subset of) a row of {\sf C}.

    \item[{\sf row\_index}]  ({\sf IN}) The index of the row in {\sf C} to 
    assign. Must be in the range $[0, \bold{nrows}({\sf C}))$.

    \item[{\sf col\_indices}]  ({\sf IN}) Pointer to the ordered set (array) of 
    indices corresponding to the locations in the specified row of {\sf C} 
    that are to be assigned.  If all elements of the row in {\sf C} are to be 
    assigned in order from index $0$ to ${\sf ncols} - 1$, then {\sf GrB\_ALL} should be 
    specified.  Regardless of execution mode and return value, this array may be
    manipulated by the caller after this operation returns without affecting any 
    deferred computations for this operation.
    If this array contains duplicate values, it implies in assignment of more 
    than one value to the same location which leads to undefined results.
    
    \item[{\sf ncols}] ({\sf IN}) The number of values in {\sf col\_indices} array.
    Must be equal to $\bold{size}({\sf u})$.

    \item[{\sf desc}] ({\sf IN}) An optional operation descriptor. If
    a \emph{default} descriptor is desired, {\sf GrB\_NULL} should be
    specified. Non-default field/value pairs are listed as follows:  \\

    \begin{tabular}{lllp{2.5in}}
        Param & Field  & Value & Description \\
        \hline
        {\sf C}    & {\sf GrB\_OUTP} & {\sf GrB\_REPLACE} &  Output row in 
        {\sf C} is cleared (all elements removed) before result is stored in it.\\

        {\sf mask} & {\sf GrB\_MASK} & {\sf GrB\_SCMP}   & Use the structural
        complement of {\sf mask}. \\
    \end{tabular}
\end{itemize}

\paragraph{Return Values}

\begin{itemize}[leftmargin=2.1in]
    \item[{\sf GrB\_SUCCESS}]         In blocking mode, the operation completed
    successfully. In non-blocking mode, this indicates that the compatibility 
    tests on dimensions and domains for the input arguments passed successfully. 
    Either way, output matrix {\sf C} is ready to be used in the next method of 
    the sequence.

    \item[{\sf GrB\_PANIC}]           Unknown internal error.

    \item[{\sf GrB\_INVALID\_OBJECT}] This is returned in any execution mode 
    whenever one of the opaque GraphBLAS objects (input or output) is in an invalid 
    state caused by a previous execution error.  Call {\sf GrB\_error()} to access 
    any error messages generated by the implementation.

    \item[{\sf GrB\_OUT\_OF\_MEMORY}] Not enough memory available for operation.

    \item[{\sf GrB\_UNINITIALIZED\_OBJECT}] One or more of the GraphBLAS objects
    has not been initialized by a call to {\sf new} (or {\sf dup} for vector or
    matrix parameters).

    \item[{\sf GrB\_INVALID\_INDEX}]    {\sf row\_index} is outside the allowable 
    range (i.e., greater than $\bold{nrows}({\sf C})$).

    \item[{\sf GrB\_INDEX\_OUT\_OF\_BOUNDS}]  A value in {\sf col\_indices} 
    is greater than or equal to $\bold{ncols}({\sf C})$.  In 
    non-blocking mode, this can be reported as an execution error.

    \item[{\sf GrB\_DIMENSION\_MISMATCH}] {\sf mask} size and number of columns
    in {\sf C} are not the same, or ${\sf ncols} \neq \bold{size}({\sf u})$.

    \item[{\sf GrB\_DOMAIN\_MISMATCH}]    The domains of the matrix and vector are
    incompatible with each other or the corresponding domains of the
    accumulation operator, or the mask's domain is not compatible with {\sf bool}.

    \item[{\sf GrB\_NULL\_POINTER}] Argument {\sf col\_indices} is a {\sf NULL} pointer.
\end{itemize}

\paragraph{Description}

This variant of {\sf GrB\_assign} computes the result of assigning a subset of
locations in a row of a GraphBLAS matrix (in a specific order) from the 
contents of a GraphBLAS vector: \\
${\sf C}({\sf row\_index},:) = {\sf u}$; or, if an 
optional binary accumulation operator ($\odot$) is provided, 
${\sf C}({\sf row\_index},:) = 
{\sf C}({\sf row\_index},:) \odot {\sf u}$. Taking order of {\sf col\_indices}
into account it is more explicitly written as:
\[
\begin{aligned}
    {\sf C}({\sf row\_index},{\sf col\_indices}[j]) = &\ {\sf u}(j),
    \ \forall \ j \ : \ 0 \leq j < {\sf ncols}, \mbox{~or~}
    \\
    {\sf C}({\sf row\_index},{\sf col\_indices}[j]) = &\ 
    {\sf C}({\sf row\_index},{\sf col\_indices}[j]) \odot {\sf u}(j),
    \ \forall \ j \ : \ 0 \leq j < {\sf ncols}
\end{aligned}
\]
Logically, this operation occurs in three steps:
\begin{enumerate}[leftmargin=0.75in]
\item[\bf Setup] The internal matrices, vectors and mask used in the computation are formed 
and their domains and dimensions are tested for compatibility.
\item[\bf Compute] The indicated computations are carried out.
\item[\bf Output] The result is written into the output matrix, possibly under 
control of a mask.
\end{enumerate}

Up to three argument vectors and matrices are used in this {\sf GrB\_assign} 
operation:
\begin{enumerate}
	\item ${\sf C} = \langle \bold{D}({\sf C}),\bold{nrows}({\sf C}),
    \bold{ncols}({\sf C}),\bold{L}({\sf C}) = \{(i,j,C_{ij}) \} \rangle$

	\item ${\sf mask} = \langle \bold{D}({\sf mask}),\bold{size}({\sf mask}),
    \bold{L}({\sf mask}) = \{(i,m_i) \} \rangle$ (optional)

	\item ${\sf u} = \langle \bold{D}({\sf u}),\bold{size}({\sf u}),
    \bold{L}({\sf u}) = \{(i,u_i) \} \rangle$
\end{enumerate}

The argument vectors, matrix, and the accumulation 
operator (if provided) are tested for domain compatibility as follows:
\begin{enumerate}
	\item The domain of {\sf mask} (if not {\sf GrB\_NULL}) must be from one of 
    the pre-defined types of Table~\ref{Tab:PredefinedTypes}.

	\item If {\sf accum} is {\sf GrB\_NULL}, then $\bold{D}({\sf C})$ must be 
    compatible with $\bold{D}({\sf u})$.

	\item If {\sf accum} is not {\sf GrB\_NULL}, then $\bold{D}({\sf C})$ must be
    compatible with $\bDin1({\sf accum})$ and $\bDout({\sf accum})$ of the accumulation operator and 
    $\bold{D}({\sf u})$ must be compatible with $\bDin2({\sf accum})$ of the accumulation operator.
\end{enumerate}
Two domains are compatible with each other if values from one domain can be cast 
to values in the other domain as per the rules of the C language.
In particular, domains from Table~\ref{Tab:PredefinedTypes} are all compatible 
with each other. A domain from a user-defined type is only compatible with itself.
If any compatibility rule above is violated, execution of {\sf GrB\_assign} ends
and the domain mismatch error listed above is returned.

The {\sf row\_index} parameter is checked for a valid value.  The following
condition must hold:
\begin{enumerate}
	\item $0\ \leq\ {\sf row\_index} \ <\ \bold{nrows}({\sf C})$
\end{enumerate}
If the rule above is violated, execution of {\sf GrB\_assign} ends 
and the invalid index error listed above is returned.

From the arguments, the internal vectors, mask, and index array used in 
the computation are formed ($\leftarrow$ denotes copy):
\begin{enumerate}
	\item The vector, $\vector{\widetilde{c}}$, is extracted from a row of {\sf C}
    as follows:
    \[
        \vector{\widetilde{c}} = \langle  \bold{D}({\sf C}), \bold{ncols}({\sf C}), 
        \{ (j, C_{ij}) \ \forall \ j : 0 \leq j < \bold{ncols}({\sf C}),
        i = {\sf row\_index}, (i, j) \in \bold{ind}({\sf C})\} \rangle
    \]

	\item One-dimensional mask, $\vector{\widetilde{m}}$, is computed from 
    argument {\sf mask} as follows:
	\begin{enumerate}
		\item	If ${\sf mask} = {\sf GrB\_NULL}$, then $\vector{\widetilde{m}} = 
        \langle \bold{ncols}({\sf C}), \{i, \forall i : 0 \leq i < 
        \bold{ncols}({\sf C}) \} \rangle$.

		\item	Otherwise, $\vector{\widetilde{m}} = 
        \langle \bold{size}({\sf mask}), \{i : i \in \bold{ind}({\sf mask}) \wedge
        ({\sf bool}){\sf mask}(i) = \true \} \rangle$.

		\item	If ${\sf desc[GrB\_MASK].GrB\_SCMP}$ is set, then 
        $\vector{\widetilde{m}} \leftarrow \neg \vector{\widetilde{m}}$.
	\end{enumerate}

	\item Vector $\vector{\widetilde{u}} \leftarrow {\sf u}$.
    
    \item The internal column index array, $\array{\widetilde{J}}$, is computed from 
    argument {\sf col\_indices} as follows:
	\begin{enumerate}
		\item	If ${\sf col\_indices} = {\sf GrB\_ALL}$, then 
        $\array{\widetilde{J}}[j] = j, \ \forall \ j : 0 \leq j < {\sf ncols}$.

		\item	Otherwise, $\array{\widetilde{J}}[j] = {\sf col\_indices}[j], 
        \ \forall \ j : 0 \leq j < {\sf ncols}$.
    \end{enumerate}
\end{enumerate}

The internal vectors, matrices, and masks are checked for dimension compatibility. 
The following conditions must hold:
\begin{enumerate}
    \item $\bold{size}(\vector{\widetilde{c}}) = \bold{size}(\vector{\widetilde{m}})$
    \item ${\sf ncols} = \bold{size}(\vector{\widetilde{u}})$.
\end{enumerate}
If any compatibility rule above is violated, execution of {\sf GrB\_assign} ends and 
the dimension mismatch error listed above is returned.

From this point forward, in {\sf GrB\_NONBLOCKING} mode, the method can 
optionally exit with {\sf GrB\_SUCCESS} return code and defer any computation 
and/or execution error codes.

We are now ready to carry out the assign and any additional 
associated operations.  We describe this in terms of two intermediate vectors:
\begin{itemize}
    \item $\vector{\widetilde{t}}$: The vector holding the elements from
    $\vector{\widetilde{u}}$ in their destination locations relative to 
    $\vector{\widetilde{c}}$.

    \item $\vector{\widetilde{z}}$: The vector holding the result after 
    application of the (optional) accumulation operator.
\end{itemize}

The intermediate vector, $\vector{\widetilde{t}}$, is created as follows:
\[
\vector{\widetilde{t}} = \langle
\bold{D}({\sf u}), \bold{size}(\vector{\widetilde{c}}),
%\bold{L}(\vector{\widetilde{t}}) =
\{(\array{\widetilde{J}}[j],\vector{\widetilde{u}}(j)) \ \forall \ j, \ 
0 \leq j < {\sf ncols} : j \in \bold{ind}(\vector{\widetilde{u}}) \} \rangle. 
\]
At this point, if any value of $\array{\widetilde{J}}[j]$ is outside the valid 
range of indices for vector $\vector{\widetilde{c}}$, computation ends and the 
method returns the index out-of-bounds error listed above. In 
{\sf GrB\_NONBLOCKING} mode, the error can be deferred until a 
sequence-terminating {\sf GrB\_wait()} is called.  Regardless, the result 
matrix, {\sf C}, is invalid from this point forward in the 
sequence.

The intermediate vector $\vector{\widetilde{z}}$ is created as follows:
\begin{itemize}
    \item If ${\sf accum} = {\sf GrB\_NULL}$, then $\vector{\widetilde{z}}$ is defined as 
    \[ 
        \vector{\widetilde{z}} =
		\langle \bold{D}({\sf C}), \bold{size}(\vector{\widetilde{c}}), 
		\{(i,z_{i}), \forall i \in (\bold{ind}(\vector{\widetilde{c}})-(\{\array{\widetilde{I}}[k],\forall k\} \cap \bold{ind}(\vector{\widetilde{c}}))) \cup 
        \bold{ind}(\vector{\widetilde{t}}) \} \rangle.
    \]
    The above expression defines the structure of vector $\vector{\widetilde{z}}$ as follows:
    We start with the structure of $\vector{\widetilde{c}}$ ($\bold{ind}(\vector{\widetilde{c}})$) and remove from 
    it all the indices of $\vector{\widetilde{c}}$ that are
    in the set of indices being assigned ($\{\array{\widetilde{I}}[k],\forall k\} \cap \bold{ind}(\vector{\widetilde{c}})$). Finally, we
    add the structure of $\vector{\widetilde{t}}$ ($\bold{ind}(\vector{\widetilde{t}})$).

    The values of the elements of $\vector{\widetilde{z}}$ are computed based on the 
    relationships between the sets of indices in $\vector{\widetilde{c}}$ 
    and $\vector{\widetilde{t}}$.
    \[
        z_{i} = \vector{\widetilde{c}}(i), \ \mbox{if}\  i \in  
        (\bold{ind}(\vector{\widetilde{c}}) - (\{\array{\widetilde{I}}[k],\forall k\}
        \cap \bold{ind}(\vector{\widetilde{c}}))),
    \]
    \[
        z_{i} = \vector{\widetilde{t}}(i), \ \mbox{if}\  i \in  
        \bold{ind}(\vector{\widetilde{t}}),
    \]
    where the difference operator refers to set difference.

    \item If ${\sf accum}$ is a binary operator, then $\vector{\widetilde{z}}$ is defined as
        \[ \langle \bDout({\sf accum}), \bold{size}(\vector{\widetilde{c}}),
        %\bold{L}(\vector{\widetilde{z}}) =
        \{(j,z_{j}) \ \forall \ j \in \bold{ind}(\vector{\widetilde{c}}) \cup 
        \bold{ind}(\vector{\widetilde{t}}) \} \rangle.\]

    The values of the elements of $\vector{\widetilde{z}}$ are computed based on the 
    relationships between the sets of indices in $\vector{\widetilde{w}}$ and 
    $\vector{\widetilde{t}}$.
\[
    z_{j} = \vector{\widetilde{c}}(j) \odot \vector{\widetilde{t}}(j), \ \mbox{if}\  
    j \in  (\bold{ind}(\vector{\widetilde{t}}) \cap \bold{ind}(\vector{\widetilde{c}})),
\]
\[
    z_{j} = \vector{\widetilde{c}}(j), \ \mbox{if}\  
    j \in  (\bold{ind}(\vector{\widetilde{c}}) - (\bold{ind}(\vector{\widetilde{t}})
    \cap \bold{ind}(\vector{\widetilde{c}}))),
\]
\[
    z_{j} = \vector{\widetilde{t}}(j), \ \mbox{if}\  j \in  
    (\bold{ind}(\vector{\widetilde{t}}) - (\bold{ind}(\vector{\widetilde{t}})
    \cap \bold{ind}(\vector{\widetilde{c}}))),
\]
where $\odot  = \bigodot({\sf accum})$, and the difference operator refers to set difference.
\end{itemize}

Finally, the set of output values that make up the $\vector{\widetilde{z}}$ 
vector are written into the column of the final result matrix, ${\sf C}({\sf row\_index},:)$.  
This is carried out under control of the mask which acts as a ``write mask''.
\begin{itemize}
\item If {\sf desc[GrB\_OUTP].GrB\_REPLACE} is set, then any values in ${\sf C}({\sf row\_index},:)$ 
on input to this operation are deleted and the new contents of the column is given by:
\[
\bold{L}({\sf C}) = \{ (i,j,C_{ij}) : i \neq {\sf row\_index} \} \cup \{({\sf row\_index},j,z_{j}) : j \in 
(\bold{ind}(\vector{\widetilde{z}}) 
\cap \bold{ind}(\vector{\widetilde{m}})) \}. 
\]

\item If {\sf desc[GrB\_OUTP].GrB\_REPLACE} is not set, the elements of 
$\vector{\widetilde{z}}$ indicated by the mask are copied into the column 
of the final result matrix, ${\sf C}({\sf row\_index},:)$, and elements of 
this column that fall outside the set indicated by the mask are unchanged:
\begin{eqnarray} 
    \bold{L}({\sf C}) & = & \{ (i,j,C_{ij}) : i \neq {\sf row\_index} \} \cup \nonumber \\
    & & \{({\sf row\_index},j,\vector{\widetilde{c}}(j)) : j \in (\bold{ind}(\vector{\widetilde{c}}) 
    \cap \bold{ind}(\neg \vector{\widetilde{m}})) \} \cup \nonumber \\
    & & \{({\sf row\_index},j,z_{j}) : j \in 
    (\bold{ind}(\vector{\widetilde{z}}) \cap \bold{ind}(\vector{\widetilde{m}})) \}. \nonumber
\end{eqnarray}
\end{itemize}

In {\sf GrB\_BLOCKING} mode, the method exits with return value 
{\sf GrB\_SUCCESS} and the new content of vector {\sf w} is as defined above
and fully computed.  
In {\sf GrB\_NONBLOCKING} mode, the method exits with return value 
{\sf GrB\_SUCCESS} and the new content of vector {\sf w} is as defined above 
but may not be fully computed; however, it can be used in the next GraphBLAS 
method call in a sequence.
