\section{Introduction}
\label{sec:intro}

Graphs are a fundamental abstraction in computer science.  They represent
relationships among objects in a finite collection.   The objects, or
\emph{vertices}, in a graph are connected by \emph{edges}, the relationships.  This leads
to the common view of a graph as two sets: a set of vertices and a set
of edges.

Graphs can also be represented as matrices.   For example, the
\emph{adjacency matrix} for a graph is constructed by labeling rows and
columns of the matrix by the vertices of the graph.  The elements of
the matrix denote the edges in the graph with matrix element $A_{ij}$
defining the edge from vertex $i$ to vertex $j$.  Most
vertices in a large graph, such as those arising in social networks,
are not connected to each other, so the matrices used for graphs tend to
be very sparse.

Many graph algorithms have been defined in the ``language of linear
algebra''~\cite{kepner2011graph}.  Mapping sparse linear algebra algorithms 
onto modern architectures is well understood, and several 
groups have built high
performance graph libraries based on sparse linear algebra~\cite{combblas,
gadepally2015graphulo, gpi2016, sundaram2015graphmat,che2016programming}.  A group
of graph algorithm researchers formed the GraphBLAS
Forum~\cite{graphblas_web} to standardize the low-level building
blocks used in these graph algorithms.  The Forum completed the
mathematical formalizations of GraphBLAS~\cite{mathgraphblas16} and
has undertaken the task to define the
binding of the C programming language onto the mathematical definition of
the GraphBLAS---the so-called GraphBLAS C API.  
 

This work has been undertaken by a subcommittee of the GraphBLAS Forum, comprised of
the authors of this paper.  The challenge in formulating
the GraphBLAS C API was to balance conflicting 
objectives: (i) simplicity and ease of use,
(ii) enabling high-performance implementations, and (iii) adherence to
the underlying mathematics. 
The C programming language has been chosen as the first target by the Forum due
to its relatively smaller feature set and its better interoperability with other high-performance 
languages such as Fortran and \Cpp.
Since this is the first specification of GraphBLAS in any language, our burden 
has largely been to define the programming concepts for the first time. We believe
our design will largely carry over to future specs of the GraphBLAS in other 
languages. 

This paper summarizes the GraphBLAS C API and the
motivation behind our decisions. 
We begin by summarizing the 
mathematical ideas behind the GraphBLAS and how those ideas
influenced our notation.  We then explain data structures,
algebraic objects, and objects that control the semantics of the functions
in the GraphBLAS C API.  We then define the
core operations in the GraphBLAS C  API and the signatures
for a subset of the functions  within the API.  We then present a betweenness centrality 
algorithm that uses the GraphBLAS C  API. We close with results
and concluding remarks.
