%-----------------------------------------------------------------------------
\subsubsection{{\sf Matrix\_assign}: Standard matrix variant}

Assign values (and implied zeros) from one GraphBLAS matrix to a subset of another GraphBLAS
matrix as specified by two sets of indices. The dimensions of the input matrix are
the same size as the row and column index arrays provided.

\paragraph{\syntax}

\begin{verbatim}
        GrB_Info GrB_Matrix_assign(GrB_Matrix              C,
                                   const GrB_Matrix        A,
                                   const GrB_Index        *row_indices,
                                   const GrB_Index         nrows,
                                   const GrB_Index        *col_indices,
                                   const GrB_Index         ncols,
                                   const GrB_Descriptor    desc);
\end{verbatim}

\paragraph{Parameters}

\begin{itemize}[leftmargin=1in]
    \item[{\sf C}]    ({\sf INOUT}) An existing GraphBLAS Matriix,
    that will be modified by the assign On output, this matrix holds
    the results of the operation.

    \item[{\sf A}]       ({\sf IN}) The GraphBLAS matrix whose contents
    are assigned to a subset of {\sf C}.

    \item[{\sf row\_indices}]  ({\sf IN}) Pointer to the ordered set
    (array) of indices corresponding to the rows of {\sf C} that are
    assigned.  If all rows of {\sf C} are to be assigned in order from
    $0$ to ${\sf nrows} - 1$, then {\sf GrB\_ALL} can be specified.
    Regardless of execution mode and return value, this array may be
    manipulated by the caller after this operation returns without
    affecting any deferred computations for this operation.  If this
    array contains duplicate values, it implies assignment of more than
    one value to the same location which leads to undefined results.

    \item[{\sf nrows}] ({\sf IN}) The number of values in {\sf
    row\_indices} array. Must be equal to $\bold{nrows}({\sf A})$ if {\sf
    A} is not tranposed, or equal to $\bold{ncols}({\sf A})$ if {\sf A}
    is transposed.

    \item[{\sf col\_indices}]  ({\sf IN}) Pointer to the ordered set
    (array) of indices corresponding to the columns of {\sf C} that
    are assigned.  If all columns of {\sf C} are to be assigned in order
    from $0$ to ${\sf ncols} - 1$, then {\sf GrB\_ALL} can be specified.
    Regardless of execution mode and return value, this array may be
    manipulated by the caller after this operation returns without
    affecting any deferred computations for this operation.  If this
    array contains duplicate values, it implies assignment of more than
    one value to the same location which leads to undefined results.

    \item[{\sf ncols}] ({\sf IN}) The number of values in {\sf
    col\_indices} array. Must be equal to $\bold{ncols}({\sf A})$ if {\sf
    A} is not tranposed, or equal to $\bold{nrows}({\sf A})$ if {\sf A}
    is transposed.

    \item[{\sf desc}] ({\sf IN}) An optional operation descriptor. If
    a \emph{default} descriptor is desired, {\sf GrB\_NULL} should be
    specified. Non-default field/value pairs are listed as follows:  \\

    \begin{tabular}{lllp{2.5in}}
        Param & Field  & Value & Description \\
        \hline
        {\sf A}    & {\sf GrB\_INP0} & {\sf GrB\_TRAN}   & Apply transpose to 
        {\sf A} before assigning elements.
    \end{tabular}
\end{itemize}

\paragraph{Return Values}

\begin{itemize}[leftmargin=2.1in]
    \item[{\sf GrB\_SUCCESS}] In blocking mode, the operation
    completed successfully. In non-blocking mode, this indicates that
    the compatibility tests on dimensions and domains for the input
    arguments passed successfully.  Either way, output matrix {\sf C}
    is ready to be used in the next method of the sequence.

    \item[{\sf GrB\_PANIC}] Unknown internal error.

    \item[{\sf GrB\_INVALID\_OBJECT}] This is returned in any execution
    mode whenever one of the opaque GraphBLAS objects (input or output)
    is in an invalid state caused by a previous execution error.
    Call {\sf GrB\_error()} to access any error messages generated by
    the implementation.

    \item[{\sf GrB\_OUT\_OF\_MEMORY}]  Not enough memory available
    for operation.

    \item[{\sf GrB\_UNINITIALIZED\_OBJECT}] One or more of the GraphBLAS
    objects has not been initialized by a call to {\sf new} (or {\sf dup}
    for matrix parameters).

    \item[{\sf GrB\_INDEX\_OUT\_OF\_BOUNDS}]  A value in {\sf
    row\_indices} is greater than or equal to $\bold{nrows}({\sf
    C})$, or a value in {\sf col\_indices} is greater than or equal to
    $\bold{ncols}({\sf C})$.  In non-blocking mode, this can be reported
    as an execution error.

    \item[{\sf GrB\_DIMENSION\_MISMATCH}] ${\sf nrows} \neq
    \bold{nrows}({\sf A})$ ($\bold{ncols}({\sf A})$ if {\sf A}
    is transposed), or ${\sf ncols} \neq \bold{ncols}({\sf A})$
    ($\bold{nrows}({\sf A})$ if {\sf A} is transposed).

    \item[{\sf GrB\_DOMAIN\_MISMATCH}]     The domains of the various
    matrices are incompatible with each other.

    \item[{\sf GrB\_NULL\_POINTER}] Either argument {\sf row\_indices}
    is a {\sf NULL} pointer, argument {\sf col\_indices} is a {\sf NULL}
    pointer, or both.
\end{itemize}

\paragraph{Description}

This variant of {\sf GrB\_assign} computes the result of assigning values of
elements form a source GraphBLAS matrix to elements of a destination GraphBLAS matrix.
More explicitly:
\[
\begin{aligned}
	{\sf C}({\sf row\_indices}[i],{\sf col\_indices}[j]) = &\ {\sf A}(i,j), \ \forall \ i,j \ : \ 0 \leq i < {\sf nrows},\ 0 \leq j < {\sf ncols}, \ \ \mbox{if {\sf A} is not transposed},
\\
	{\sf C}({\sf row\_indices}[i],{\sf col\_indices}[j]) = &\ {\sf A}(j,i), \ \forall \ i,j \ : \ 0 \leq i < {\sf ncols},\ 0 \leq j < {\sf nrows}, \ \ \mbox{if {\sf A} is transposed}.
\end{aligned}
\]  
If a particular element ${\sf A}(i,j)$ does not exist, the corresponding
element of {\sf C} is also removed. 
Elements of {\sf C} outside the ranges of $({\sf row\_indices}[i])$ and $({\sf col\_indics}[j])$ are not affected.

Logically, this operation occurs in three steps:
\begin{enumerate}[leftmargin=0.75in]
\item[Setup] The internal matrices used in the computation are formed 
and their domains and dimensions are tested for compatibility.
\item[Compute] The indicated computations are carried out.
\item[Output] The result is written into the output matrix.
\end{enumerate}

Two argument matrices are used in the {\sf GrB\_assign} operation:
\begin{enumerate}
	\item ${\sf C} = \langle \bold{D}({\sf C}),\bold{nrows}({\sf C}),
    \bold{ncols}({\sf C}),\bold{L}({\sf C}) = \{(i,j,C_{ij}) \} \rangle$
    
	\item ${\sf A} = \langle \bold{D}({\sf A}),\bold{nrows}({\sf A}), 
    \bold{ncols}({\sf A}),\bold{L}({\sf A}) = \{(i,j,A_{ij}) \} \rangle$
\end{enumerate}

The argument matrices 
are tested for domain compatibility as follows:
\begin{enumerate}
	\item $\bold{D}({\sf C})$ must be compatible with $\bold{D}({\sf A})$.
\end{enumerate}
Two domains are compatible with each other if values from one domain can be cast 
to values in the other domain as per the rules of the C language.
In particular, domains from Table~\ref{Tab:PredefinedTypes} are all compatible 
with each other. A domain from a user-defined type is only compatible with itself.
If any compatibility rule above is violated, execution of {\sf GrB\_assign} ends
and the domain mismatch error listed above is returned.

From the arguments, the internal matrices and index arrays used in 
the computation are formed ($\leftarrow$ denotes copy):
\begin{enumerate}
	\item Matrix $\matrix{\widetilde{C}} \leftarrow {\sf C}$.

	\item Matrix $\matrix{\widetilde{A}} \leftarrow 
    {\sf desc[GrB\_INP0].GrB\_TRAN} \ ? \ {\sf A}^T : {\sf A}$.

    \item The internal row index array, $\grbarray{\widetilde{I}}$, is computed from 
    argument {\sf row\_indices} as follows:
	\begin{enumerate}
		\item	If ${\sf row\_indices} = {\sf GrB\_ALL}$, then 
        $\grbarray{\widetilde{I}}[i] = i, \forall i : 0 \leq i < {\sf nrows}$.

		\item	Otherwise, $\grbarray{\widetilde{I}}[i] = {\sf row\_indices}[i], 
        \forall i : 0 \leq i < {\sf nrows}$.
    \end{enumerate}
    
    \item The internal column index array, $\grbarray{\widetilde{J}}$, is computed from 
    argument {\sf col\_indices} as follows:
	\begin{enumerate}
		\item	If ${\sf col\_indices} = {\sf GrB\_ALL}$, then 
        $\grbarray{\widetilde{J}}[j] = j, \forall j : 0 \leq j < {\sf ncols}$.

		\item	Otherwise, $\grbarray{\widetilde{J}}[j] = {\sf col\_indices}[j], 
        \forall j : 0 \leq j < {\sf ncols}$.
    \end{enumerate}
\end{enumerate}

The internal matrices and mask are checked for shape compatibility. The following 
conditions must hold:
\begin{enumerate}
    \item $\bold{nrows}(\matrix{\widetilde{A}}) = {\sf nrows}$.

    \item $\bold{ncols}(\matrix{\widetilde{A}}) = {\sf ncols}$.
\end{enumerate}
If any compatibility rule above is violated, execution of {\sf GrB\_assign} ends and 
the dimension mismatch error listed above is returned.

From this point forward, in {\sf GrB\_NONBLOCKING} mode, the method can 
optionally exit with {\sf GrB\_SUCCESS} return code and defer any computation 
and/or execution error codes.

We are now ready to carry out the assign and any additional 
associated operations.  We describe this in terms of an intermediate matrix:
\begin{itemize}
	\item $\matrix{\widetilde{T}}$: The matrix holding the contents from
    $\matrix{\widetilde{A}}$ in their destination locations relative to
    $\matrix{\widetilde{C}}$.
\end{itemize}

The intermediate matrix, $\matrix{\widetilde{T}}$, is created as follows:
\[ 
\begin{aligned}
\matrix{\widetilde{T}} = \langle & \bold{D}({\sf A}), 
                           \bold{nrows}(\matrix{\widetilde{C}}), 
                           \bold{ncols}(\matrix{\widetilde{C}}), \\
 & \{ (\grbarray{\widetilde{I}}[i],\grbarray{\widetilde{J}}[j],\matrix{\widetilde{A}}(i,j)) 
\ \forall \ (i,j), \ 0 \leq i < {\sf nrows}, \ 0 \leq j < {\sf ncols} :
(i,j) \in \bold{ind}(\matrix{\widetilde{A}}) \} \rangle. 
\end{aligned}
\]
At this point, if any value in the $\grbarray{\widetilde{I}}$ array is not in
the range $[0,\ \bold{nrows}(\matrix{\widetilde{C}}) )$ or any value in the 
$\grbarray{\widetilde{J}}$ array is not in the range 
$[0,\ \bold{ncols}(\matrix{\widetilde{C}}))$, the execution of {\sf GrB\_assign} 
ends and the index out-of-bounds error listed above is generated.  In 
{\sf GrB\_NONBLOCKING} mode, the error can be deferred until a 
sequence-terminating {\sf GrB\_wait()} is called.  Regardless, the result 
matrix {\sf C} is invalid from this point forward in the sequence.

Finally, the assign operation is completed by updating
the contents of matrix {\sf C} according to the following expression:
\[
	\bold{L}({\sf C}) = (\bold{L}({\matrix{\widetilde{C}}}) - \{(\grbarray{\widetilde{I}}[i],\grbarray{\widetilde{J}}[j],C) \forall i,j : (\grbarray{\widetilde{I}}[i],\grbarray{\widetilde{J}}[j]) \in \bold{ind}({\sf C})\})
	\cup \{(i,j,\matrix{\widetilde{T}}(i,j)) \forall (i,j) \in \bold{ind}({\widetilde{T}})\},
\]
where the difference operator in the previous expressions refers to set difference.
That is, first existing elements of {\sf C} with row indices in $\grbarray{\widetilde{I}}$ and column indices in $\grbarray{\widetilde{J}}$ are removed,
and then the elements of $\widetilde{T}$ are added in the right place.

In {\sf GrB\_BLOCKING} mode, the method exits with return value 
{\sf GrB\_SUCCESS} and the new content of matrix {\sf C} is as defined above
and fully computed.  
In {\sf GrB\_NONBLOCKING} mode, the method exits with return value 
{\sf GrB\_SUCCESS} and the new content of matrix {\sf C} is as defined above 
but may not be fully computed; however, it can be used in the next GraphBLAS 
method call in a sequence.

%-----------------------------------------------------------------------------
\subsubsection{{\sf Matrix\_assign}: Column variant}

Assign the contents a vector to a subset of elements in one column of a matrix. 
Note that since the output cannot be transposed, a different variant of
{\sf assign} is provided to assign to a row of a matrix.

\paragraph{\syntax}

\begin{verbatim}
        GrB_Info GrB_Matrix_assign(GrB_Matrix              C,
                                   const GrB_Vector        mask,
                                   const GrB_BinaryOp      accum,
                                   const GrB_Vector        u,
                                   const GrB_Index        *row_indices,
                                   const GrB_Index         nrows,
                                   GrB_Index               col_index,
                                   const GrB_Descriptor    desc); 
\end{verbatim}

\paragraph{Parameters}

\begin{itemize}[leftmargin=1.1in]
    \item[{\sf C}]    ({\sf INOUT}) An existing GraphBLAS Matrix.  On input,
    the matrix provides values that may be accumulated with the result of the
    assign operation.  On output, this matrix holds the results of the
    operation.

    \item[{\sf mask}] ({\sf IN}) An optional ``write'' mask that controls which
    results from this operation are stored into the specified column of the output matrix {\sf C}. The 
    mask dimensions must match those of a single column of the matrix {\sf C} and the domain of the 
    {\sf Mask} matrix must be of type {\sf bool} or any of the predefined 
    ``built-in'' types in Table~\ref{Tab:PredefinedTypes}.  If the default
    vector is desired (\ie, with correct dimensions and filled with {\sf true}), 
    {\sf GrB\_NULL} should be specified.

    \item[{\sf accum}]    ({\sf IN}) An optional operator used for accumulating
    entries into existing {\sf C} entries: ${\sf accum} = \langle D_x,
    D_y, D_z,\odot \rangle$. If assignment rather than accumulation is
    desired, {\sf GrB\_NULL} should be specified.

    \item[{\sf u}]       ({\sf IN}) The GraphBLAS vector whose contents are 
    assigned to (a subset of) a column of {\sf C}.

    \item[{\sf row\_indices}]  ({\sf IN}) Pointer to the ordered set (array) of 
    indices corresponding to the locations in the specified column of {\sf C} 
    that are to be assigned.  If all elements of the column in {\sf C} are to be 
    assigned in order from index $0$ to ${\sf nrows} - 1$, then {\sf GrB\_ALL} should be 
    specified.  Regardless of execution mode and return value, this array may be
    manipulated by the caller after this operation returns without affecting any 
    deferred computations for this operation.
    If this array contains duplicate values, it implies in assignment of more 
    than one value to the same location which leads to undefined results.
    
    \item[{\sf nrows}] ({\sf IN}) The number of values in {\sf row\_indices} array.
    Must be equal to $\bold{size}({\sf u})$.
    
    \item[{\sf col\_index}]  ({\sf IN}) The index of the column in {\sf C} to 
    assign. Must be in the range $[0, \bold{ncols}({\sf C}))$.

    \item[{\sf desc}] ({\sf IN}) An optional operation descriptor. If
    a \emph{default} descriptor is desired, {\sf GrB\_NULL} should be
    specified. Non-default field/value pairs are listed as follows:  \\

    \begin{tabular}{lllp{2.5in}}
        Param & Field  & Value & Description \\
        \hline
        {\sf C}    & {\sf GrB\_OUTP} & {\sf GrB\_REPLACE} &  Output column in 
        {\sf C} is cleared (all elements removed) before result is stored in it.\\
    
        {\sf mask} & {\sf GrB\_MASK} & {\sf GrB\_SCMP}   & Use the structural 
        complement of {\sf mask}. \\
    \end{tabular}

\end{itemize}

\paragraph{Return Values}

\begin{itemize}[leftmargin=2.1in]
    \item[{\sf GrB\_SUCCESS}]         In blocking mode, the operation completed
    successfully. In non-blocking mode, this indicates that the compatibility 
    tests on dimensions and domains for the input arguments passed successfully. 
    Either way, output matrix {\sf C} is ready to be used in the next method of 
    the sequence.

    \item[{\sf GrB\_PANIC}]            Unknown internal error.
    
    \item[{\sf GrB\_INVALID\_OBJECT}] This is returned in any execution mode 
    whenever one of the opaque GraphBLAS objects (input or output) is in an invalid 
    state caused by a previous execution error.  Call {\sf GrB\_error()} to access 
    any error messages generated by the implementation.

    \item[{\sf GrB\_OUT\_OF\_MEMORY}]  Not enough memory available for operation.
    
    \item[{\sf GrB\_UNINITIALIZED\_OBJECT}] One or more of the GraphBLAS objects
    has not been initialized by a call to {\sf new} (or {\sf dup} for vector or
    matrix parameters).

    \item[{\sf GrB\_INVALID\_INDEX}]    {\sf col\_index} is outside the allowable 
    range (i.e., greater than $\bold{ncols}({\sf C})$).

    \item[{\sf GrB\_INDEX\_OUT\_OF\_BOUNDS}]  A value in {\sf row\_indices} 
    is greater than or equal to $\bold{nrows}({\sf C})$.  In 
    non-blocking mode, this can be reported as an execution error.
    
    \item[{\sf GrB\_DIMENSION\_MISMATCH}] {\sf mask} size and number of rows
    in {\sf C} are not the same, or ${\sf nrows} \neq \bold{size}({\sf u})$.

    \item[{\sf GrB\_DOMAIN\_MISMATCH}]     The domains of the matrix and vector
    are incompatible with each other or the corresponding domains of the 
    accumulation operator, or the mask's domain is not compatible with {\sf bool}.

    \item[{\sf GrB\_NULL\_POINTER}] Argument {\sf row\_indices} is a {\sf NULL} pointer.
\end{itemize}

\paragraph{Description}

This variant of {\sf GrB\_assign} computes the result of assigning a subset of
locations in a column of a GraphBLAS matrix (in a specific order) from the 
contents of a GraphBLAS vector: \\
${\sf C}(:,{\sf col\_index}) = {\sf u}$; or, if an 
optional binary accumulation operator ($\odot$) is provided, 
${\sf C}(:,{\sf col\_index}) = 
{\sf C}(:,{\sf col\_index}) \odot {\sf u}$. Taking order of {\sf row\_indices} 
into account, it is more explicitly written as:
\[
\begin{aligned}
    {\sf C}({\sf row\_indices}[i],{\sf col\_index}) = &\ {\sf u}(i) 
    \ \forall \ i \ : \ 0 \leq i < {\sf nrows}, \mbox{~or~}
    \\
    {\sf C}({\sf row\_indices}[i],{\sf col\_index}) = &\ 
    {\sf C}({\sf row\_indices}[i],{\sf col\_index}) \odot {\sf u}(i) 
    \ \forall \ i \ : \ 0 \leq i < {\sf nrows}
\end{aligned}
\]  
Logically, this operation occurs in three steps:
\begin{enumerate}[leftmargin=0.75in]
\item[\bf Setup] The internal matrices, vectors and mask used in the computation are formed 
and their domains and dimensions are tested for compatibility.
\item[\bf Compute] The indicated computations are carried out.
\item[\bf Output] The result is written into the output matrix, possibly under 
control of a mask.
\end{enumerate}

Up to three argument vectors and matrices are used in this {\sf GrB\_assign} 
operation:
\begin{enumerate}
	\item ${\sf C} = \langle \bold{D}({\sf C}),\bold{nrows}({\sf C}),
    \bold{ncols}({\sf C}),\bold{L}({\sf C}) = \{(i,j,C_{ij}) \} \rangle$
    
	\item ${\sf mask} = \langle \bold{D}({\sf mask}),\bold{size}({\sf mask}),
    \bold{L}({\sf mask}) = \{(i,m_{i}) \} \rangle$ (optional)

	\item ${\sf u} = \langle \bold{D}({\sf u}),\bold{size}({\sf u}),
    \bold{L}({\sf u}) = \{(i,u_i) \} \rangle$
\end{enumerate}

The argument vectors, matrix, and the accumulation 
operator (if provided) are tested for domain compatibility as follows:
\begin{enumerate}
	\item The domain of {\sf mask} (if not {\sf GrB\_NULL}) must be from one of 
    the pre-defined types of Table~\ref{Tab:PredefinedTypes}.

	\item If {\sf accum} is {\sf GrB\_NULL}, then $\bold{D}({\sf C})$ must be 
    compatible with $\bold{D}({\sf u})$.

	\item If {\sf accum} is not {\sf GrB\_NULL}, then $\bold{D}({\sf C})$ must be
    compatible with $D_x$ and $D_z$ of the accumulation operator and 
    $\bold{D}({\sf u})$ must be compatible with $D_y$ of the accumulation operator.
\end{enumerate}
Two domains are compatible with each other if values from one domain can be cast 
to values in the other domain as per the rules of the C language.
In particular, domains from Table~\ref{Tab:PredefinedTypes} are all compatible 
with each other. A domain from a user-defined type is only compatible with itself.
If any compatibility rule above is violated, execution of {\sf GrB\_assign} ends
and the domain mismatch error listed above is returned.

The {\sf col\_index} parameter is checked for a valid value.  The following
condition must hold:
\begin{enumerate}
	\item $0\ \leq\ {\sf col\_index} \ <\ \bold{ncols}({\sf C})$
\end{enumerate}
If the rule above is violated, execution of {\sf GrB\_assign} ends 
and the invalid index error listed above is returned.

From the arguments, the internal vectors, mask, and index array used in 
the computation are formed ($\leftarrow$ denotes copy):
\begin{enumerate}
	\item The vector, $\vector{\widetilde{c}}$, is extracted from a column of {\sf C}
    as follows:
    \[
        \vector{\widetilde{c}} = \langle  \bold{D}({\sf C}), \bold{nrows}({\sf C}), 
        \{ (i, C_{ij}) \ \forall \ i : 0 \leq i < \bold{nrows}({\sf C}),
        j = {\sf col\_index}, (i, j) \in \bold{ind}({\sf C})\} \rangle
    \]

	\item One-dimensional mask, $\vector{\widetilde{m}}$, is computed from 
    argument {\sf mask} as follows:
	\begin{enumerate}
		\item	If ${\sf mask} = {\sf GrB\_NULL}$, then $\vector{\widetilde{m}} = 
        \langle \bold{nrows}({\sf C}), \{i,\ \forall \ i : 0 \leq i < 
        \bold{nrows}({\sf C}) \} \rangle$.

		\item	Otherwise, $\vector{\widetilde{m}} = 
        \langle \bold{size}({\sf mask}), \{i : i \in \bold{ind}({\sf mask}) \wedge
        ({\sf bool}){\sf mask}(i) = \true \} \rangle$.

		\item	If ${\sf desc[GrB\_MASK].GrB\_SCMP}$ is \true, then 
        $\vector{\widetilde{m}} \leftarrow \neg \vector{\widetilde{m}}$.
	\end{enumerate}

	\item Vector $\vector{\widetilde{u}} \leftarrow {\sf u}$.
    
    \item The internal row index array, $\grbarray{\widetilde{I}}$, is computed from 
    argument {\sf row\_indices} as follows:
	\begin{enumerate}
		\item	If ${\sf row\_indices} = {\sf GrB\_ALL}$, then 
        $\grbarray{\widetilde{I}}[i] = i, \ \forall \ i : 0 \leq i < {\sf nrows}$.

		\item	Otherwise, $\grbarray{\widetilde{I}}[i] = {\sf row\_indices}[i], 
        \forall i : 0 \leq i < {\sf nrows}$.
    \end{enumerate}
\end{enumerate}

The internal vectors, matrices, and masks are checked for dimension compatibility. 
The following conditions must hold:
\begin{enumerate}
	\item $\bold{size}(\vector{\widetilde{c}}) = \bold{size}(\vector{\widetilde{m}})$
    \item ${\sf nrows} = \bold{size}(\vector{\widetilde{u}})$.
\end{enumerate}
If any compatibility rule above is violated, execution of {\sf GrB\_assign} ends and 
the dimension mismatch error listed above is returned.

From this point forward, in {\sf GrB\_NONBLOCKING} mode, the method can 
optionally exit with {\sf GrB\_SUCCESS} return code and defer any computation 
and/or execution error codes.

We are now ready to carry out the assign and any additional 
associated operations.  We describe this in terms of two intermediate vectors:
\begin{itemize}
    \item $\vector{\widetilde{t}}$: The vector holding the elements from
    $\vector{\widetilde{u}}$ in their destination locations relative to 
    $\vector{\widetilde{c}}$.
    
    \item $\vector{\widetilde{z}}$: The vector holding the result after 
    application of the (optional) accumulation.
\end{itemize}

The intermediate vector, $\vector{\widetilde{t}}$, is created as follows:
\[
\vector{\widetilde{t}} = \langle
\bold{D}({\sf u}), \bold{size}(\vector{\widetilde{c}}),
%\bold{L}(\vector{\widetilde{t}}) =
\{(\grbarray{\widetilde{I}}[i],\vector{\widetilde{u}}(i))\ \forall \ i,
 0 \leq i < {\sf nrows} : i \in \bold{ind}(\vector{\widetilde{u}}) \} \rangle. 
\]
At this point, if any value of $\grbarray{\widetilde{I}}[i]$ is outside the valid 
range of indices for vector $\vector{\widetilde{c}}$, computation ends and the 
method returns the index out-of-bounds error listed above. In 
{\sf GrB\_NONBLOCKING} mode, the error can be deferred until a 
sequence-terminating {\sf GrB\_wait()} is called.  Regardless, the result 
matrix, {\sf C}, is invalid from this point forward in the 
sequence.

The intermediate vector $\vector{\widetilde{z}}$ is created as follows:
\begin{itemize}
    \item If ${\sf accum} = {\sf GrB\_NULL}$, then 
    $\vector{\widetilde{z}} = \vector{\widetilde{t}}$.

    \item If ${\sf accum} = \langle D_x, D_y, D_z, \odot \rangle$, then vector 
    $\vector{\widetilde{z}}$ is defined as 
        \[ 
        \vector{\widetilde{z}} = 
        \langle D_z, \bold{size}(\vector{\widetilde{c}}), 
        %\bold{L}(\vector{\widetilde{z}}) = 
        \{(i,z_{i}),   
        \ \forall \ i \in \bold{ind}(\vector{\widetilde{c}}) \cup 
        \bold{ind}(\vector{\widetilde{t}}) \} \rangle.\]
    The values of the elements of $\vector{\widetilde{z}}$ are computed based on the relationships between the sets of indices in $\vector{\widetilde{w}}$ and $\vector{\widetilde{t}}$.
\[
z_{i} = \vector{\widetilde{c}}(i) \odot \vector{\widetilde{t}}(i), \ \mbox{if}\  
i \in  (\bold{ind}(\vector{\widetilde{t}}) \cap \bold{ind}(\vector{\widetilde{c}})),
\]
\[
z_{i} = \vector{\widetilde{c}}(i), \ \mbox{if}\  i \in  (\bold{ind}(
\vector{\widetilde{c}}) - (\bold{ind}(\vector{\widetilde{t}}) \cap 
\bold{ind}(\vector{\widetilde{c}}))),
\]
\[
z_{i} = \vector{\widetilde{t}}(i), \ \mbox{if}\  i \in  (\bold{ind}(
\vector{\widetilde{t}}) - (\bold{ind}(\vector{\widetilde{t}}) \cap 
\bold{ind}(\vector{\widetilde{c}}))).
\]
where the difference operator in the previous expressions refers to set difference.
\end{itemize}

Finally, the set of output values that make up the $\vector{\widetilde{z}}$ 
vector are written into the column of the final result matrix, 
${\sf C}(:,{\sf col\_index})$.  This is carried out under control of the mask 
which acts as a ``write mask''.
\begin{itemize}
    \item If {\sf desc[GrB\_OUTP].GrB\_REPLACE} is set then any values in 
    ${\sf C}(:,{\sf col\_index})$ on input to {\sf GrB\_assign()} are deleted 
    and the new content of the column is given by:
    \[ 
		\bold{L}({\sf C}) = \{ (i,j,C_{ij}) : j \neq {\sf col\_index} \} \cup \{(i,{\sf col\_index},z_{i}) : i \in 
    (\bold{ind}(\vector{\widetilde{z}}) \cap \bold{ind}(\vector{\widetilde{m}})) \}. 
    \]

    \item If {\sf desc[GrB\_OUTP].GrB\_REPLACE} is not set, the elements of 
    $\vector{\widetilde{z}}$ indicated by the mask are copied into the column 
    of the final result matrix, ${\sf C}(:,{\sf col\_index})$, and elements of 
    this column that fall outside the set indicated by the mask are unchanged:
		\begin{eqnarray} 
			\bold{L}({\sf C}) & = & \{ (i,j,C_{ij}) : j \neq {\sf col\_index} \} \cup \nonumber \\
			& & \{(i,{\sf col\_index},\vector{\widetilde{c}}(i)) : i \in (\bold{ind}(\vector{\widetilde{c}}) 
		\cap \bold{ind}(\neg \vector{\widetilde{m}})) \} \cup \nonumber \\
			& & \{(i,{\sf col\_index},z_{i}) : i \in 
    (\bold{ind}(\vector{\widetilde{z}}) \cap \bold{ind}(\vector{\widetilde{m}})) \}. \nonumber
		\end{eqnarray}
\end{itemize}

In {\sf GrB\_BLOCKING} mode, the method exits with return value 
{\sf GrB\_SUCCESS} and the new content of vector {\sf w} is as defined above
and fully computed.  
In {\sf GrB\_NONBLOCKING} mode, the method exits with return value 
{\sf GrB\_SUCCESS} and the new content of vector {\sf w} is as defined above 
but may not be fully computed; however, it can be used in the next GraphBLAS 
method call in a sequence.


%-----------------------------------------------------------------------------
\subsubsection{{\sf Matrix\_assign}: Row variant}

Assign the contents a vector to a subset of elements in one row of a matrix. 
Note that since the output cannot be transposed, a different variant of
{\sf assign} is provided to assign to a column of a matrix.

\paragraph{\syntax}

\begin{verbatim}
        GrB_Info GrB_Matrix_assign(GrB_Matrix              C,
                                   const GrB_Vector        mask,
                                   const GrB_BinaryOp      accum,
                                   const GrB_Vector        u,
                                   GrB_Index               row_index,
                                   const GrB_Index        *col_indices,
                                   const GrB_Index         ncols,
                                   const GrB_Descriptor    desc); 
\end{verbatim}

\paragraph{Parameters}

\begin{itemize}[leftmargin=1.1in]
    \item[{\sf C}]    ({\sf INOUT}) An existing GraphBLAS Matrix.  On input,
    the matrix provides values that may be accumulated with the result of the
    assign operation.  On output, this matrix holds the results of the
    operation.

    \item[{\sf mask}] ({\sf IN}) An optional ``write'' mask that controls which
    results from this operation are stored into the specified row of the output matrix {\sf C}. The 
    mask dimensions must match those of a single row of the matrix {\sf C} and the domain of the 
    {\sf Mask} matrix must be of type {\sf bool} or any of the predefined 
    ``built-in'' types in Table~\ref{Tab:PredefinedTypes}.  If the default
    vector is desired (\ie, with correct dimensions and filled with {\sf true}), 
    {\sf GrB\_NULL} should be specified.
    
    \item[{\sf accum}]    ({\sf IN}) An optional operator used for accumulating
    entries into existing {\sf C} entries: ${\sf accum} = \langle D_x,
    D_y, D_z,\odot \rangle$. If assignment rather than accumulation is
    desired, {\sf GrB\_NULL} should be specified.

    \item[{\sf u}]       ({\sf IN}) The GraphBLAS vector whose contents are 
    assigned to (a subset of) a row of {\sf C}.

    \item[{\sf row\_index}]  ({\sf IN}) The index of the row in {\sf C} to 
    assign. Must be in the range $[0, \bold{nrows}({\sf C}))$.

    \item[{\sf col\_indices}]  ({\sf IN}) Pointer to the ordered set (array) of 
    indices corresponding to the locations in the specified row of {\sf C} 
    that are to be assigned.  If all elements of the row in {\sf C} are to be 
    assigned in order from index $0$ to ${\sf ncols} - 1$, then {\sf GrB\_ALL} should be 
    specified.  Regardless of execution mode and return value, this array may be
    manipulated by the caller after this operation returns without affecting any 
    deferred computations for this operation.
    If this array contains duplicate values, it implies in assignment of more 
    than one value to the same location which leads to undefined results.
    
    \item[{\sf ncols}] ({\sf IN}) The number of values in {\sf col\_indices} array.
    Must be equal to $\bold{size}({\sf u})$.

    \item[{\sf desc}] ({\sf IN}) An optional operation descriptor. If
    a \emph{default} descriptor is desired, {\sf GrB\_NULL} should be
    specified. Non-default field/value pairs are listed as follows:  \\

    \begin{tabular}{lllp{2.5in}}
        Param & Field  & Value & Description \\
        \hline
        {\sf C}    & {\sf GrB\_OUTP} & {\sf GrB\_REPLACE} &   Output row in 
        {\sf C} is cleared (all elements removed) before result is stored in it. \\
        
        {\sf mask} & {\sf GrB\_MASK} & {\sf GrB\_SCMP} & Use the structural 
        complement of {\sf mask}. \\
    \end{tabular}

\end{itemize}

\paragraph{Return Values}

\begin{itemize}[leftmargin=2.1in]
    \item[{\sf GrB\_SUCCESS}]         In blocking mode, the operation completed
    successfully. In non-blocking mode, this indicates that the compatibility 
    tests on dimensions and domains for the input arguments passed successfully. 
    Either way, output matrix {\sf C} is ready to be used in the next method of 
    the sequence.

    \item[{\sf GrB\_PANIC}]            Unknown internal error.

    \item[{\sf GrB\_INVALID\_OBJECT}] This is returned in any execution mode 
    whenever one of the opaque GraphBLAS objects (input or output) is in an invalid 
    state caused by a previous execution error.  Call {\sf GrB\_error()} to access 
    any error messages generated by the implementation.

    \item[{\sf GrB\_OUT\_OF\_MEMORY}]  Not enough memory available for operation.

    \item[{\sf GrB\_UNINITIALIZED\_OBJECT}] One or more of the GraphBLAS objects
    has not been initialized by a call to {\sf new} (or {\sf dup} for vector or
    matrix parameters).

    \item[{\sf GrB\_INVALID\_INDEX}]    {\sf row\_index} is outside the allowable 
    range (i.e., greater than $\bold{nrows}({\sf C})$).

    \item[{\sf GrB\_INDEX\_OUT\_OF\_BOUNDS}]  A value in {\sf col\_indices} 
    is greater than or equal to $\bold{ncols}({\sf C})$.  In 
    non-blocking mode, this can be reported as an execution error.

    \item[{\sf GrB\_DIMENSION\_MISMATCH}] {\sf mask} size and number of columns
    in {\sf C} are not the same, or ${\sf ncols} \neq \bold{size}({\sf u})$.

    \item[{\sf GrB\_DOMAIN\_MISMATCH}]     The domains of the matrix and vector
    are incompatible with each other or the corresponding domains of the 
    accumulation operator, or the mask's domain is not compatible with {\sf bool}.

    \item[{\sf GrB\_NULL\_POINTER}] Argument {\sf col\_indices} is a {\sf NULL} pointer.
\end{itemize}

\paragraph{Description}

This variant of {\sf GrB\_assign} computes the result of assigning a subset of
locations in a row of a GraphBLAS matrix (in a specific order) from the 
contents of a GraphBLAS vector: 
${\sf C}({\sf row\_index},:) = {\sf u}$; or, if an 
optional binary accumulation operator ($\odot$) is provided, 
${\sf C}({\sf row\_index},:) = 
{\sf C}({\sf row\_index},:) \odot {\sf u}$. Taking order into account it is
more explicitly written as:
\[
\begin{aligned}
    {\sf C}({\sf row\_index},{\sf col\_indices}[j]) = &\ {\sf u}(j),
    \ \forall \ j \ : \ 0 \leq j < {\sf ncols}, \mbox{~or~}
    \\
    {\sf C}({\sf row\_index},{\sf col\_indices}[j]) = &\ {\sf C}({\sf row\_index},{\sf col\_indices}[j]) \odot {\sf u}(j), 
    \ \forall \ j \ : \ 0 \leq j < {\sf ncols}
\end{aligned}
\]  
Logically, this operation occurs in three steps:
\begin{enumerate}[leftmargin=0.75in]
\item[\bf Setup] The internal matrices, vectors and mask used in the computation are formed 
and their domains and dimensions are tested for compatibility.
\item[\bf Compute] The indicated computations are carried out.
\item[\bf Output] The result is written into the output matrix, possibly under 
control of a mask.
\end{enumerate}

Up to three argument vectors and matrices are used in this {\sf GrB\_assign} 
operation:
\begin{enumerate}
	\item ${\sf C} = \langle \bold{D}({\sf C}),\bold{nrows}({\sf C}),
    \bold{ncols}({\sf C}),\bold{L}({\sf C}) = \{(i,j,C_{ij}) \} \rangle$
    
	\item ${\sf mask} = \langle \bold{D}({\sf mask}),\bold{size}({\sf mask}),
    \bold{L}({\sf mask}) = \{(i,m_{i}) \} \rangle$ (optional)

	\item ${\sf u} = \langle \bold{D}({\sf u}),\bold{size}({\sf u}),
    \bold{L}({\sf u}) = \{(i,u_i) \} \rangle$
\end{enumerate}

The argument vectors, matrix, and the accumulation 
operator (if provided) are tested for domain compatibility as follows:
\begin{enumerate}
	\item The domain of {\sf mask} (if not {\sf GrB\_NULL}) must be from one of 
    the pre-defined types of Table~\ref{Tab:PredefinedTypes}.

	\item If {\sf accum} is {\sf GrB\_NULL}, then $\bold{D}({\sf C})$ must be 
    compatible with $\bold{D}({\sf u})$.

	\item If {\sf accum} is not {\sf GrB\_NULL}, then $\bold{D}({\sf C})$ must be
    compatible with $D_x$ and $D_z$ of the accumulation operator and 
    $\bold{D}({\sf u})$ must be compatible with $D_y$ of the accumulation operator.
\end{enumerate}
Two domains are compatible with each other if values from one domain can be cast 
to values in the other domain as per the rules of the C language.
In particular, domains from Table~\ref{Tab:PredefinedTypes} are all compatible 
with each other. A domain from a user-defined type is only compatible with itself.
If any compatibility rule above is violated, execution of {\sf GrB\_assign} ends
and the domain mismatch error listed above is returned.

The {\sf row\_index} parameter is checked for a valid value.  The following
condition must hold:
\begin{enumerate}
	\item $0\ \leq\ {\sf row\_index} \ <\ \bold{nrows}({\sf C})$
\end{enumerate}
If the rule above is violated, execution of {\sf GrB\_assign} ends 
and the invalid index error listed above is returned.

From the arguments, the internal vectors, mask, and index array used in 
the computation are formed ($\leftarrow$ denotes copy):
\begin{enumerate}
	\item The vector, $\vector{\widetilde{c}}$, is extracted from a row of {\sf C}
    as follows:
    \[
        \vector{\widetilde{c}} = \langle  \bold{D}({\sf C}), \bold{ncols}({\sf C}), 
        \{ (j, C_{ij}) \ \forall \ j : 0 \leq j < \bold{ncols}({\sf C}),
        i = {\sf row\_index}, (i, j) \in \bold{ind}({\sf C})\} \rangle
    \]

	\item One-dimensional mask, $\vector{\widetilde{m}}$, is computed from 
    argument {\sf mask} as follows:
	\begin{enumerate}
		\item	If ${\sf mask} = {\sf GrB\_NULL}$, then $\vector{\widetilde{m}} = 
        \langle \bold{ncols}({\sf C}), \{i, \forall i : 0 \leq i < 
        \bold{ncols}({\sf C}) \} \rangle$.

		\item	Otherwise, $\vector{\widetilde{m}} = 
        \langle \bold{size}({\sf mask}), \{i : i \in \bold{ind}({\sf mask}) \wedge
        ({\sf bool}){\sf mask}(i) = \true \} \rangle$.

		\item	If ${\sf desc[GrB\_MASK].GrB\_SCMP}$ is \true, then 
        $\vector{\widetilde{m}} \leftarrow \neg \vector{\widetilde{m}}$.
	\end{enumerate}

	\item Vector $\vector{\widetilde{u}} \leftarrow {\sf u}$.
    
    \item The internal column index array, $\grbarray{\widetilde{J}}$, is computed from 
    argument {\sf col\_indices} as follows:
	\begin{enumerate}
		\item	If ${\sf col\_indices} = {\sf GrB\_ALL}$, then 
        $\grbarray{\widetilde{J}}[j] = j, \forall j : 0 \leq j < {\sf ncols}$.

		\item	Otherwise, $\grbarray{\widetilde{J}}[j] = {\sf col\_indices}[j], 
        \forall j : 0 \leq j < {\sf ncols}$.
    \end{enumerate}
\end{enumerate}

The internal vectors, matrices, and masks are checked for dimension compatibility. 
The following conditions must hold:
\begin{enumerate}
	\item $\bold{size}(\vector{\widetilde{c}}) = \bold{size}(\vector{\widetilde{m}})$
    \item ${\sf ncols} = \bold{size}(\vector{\widetilde{u}})$.
\end{enumerate}
If any compatibility rule above is violated, execution of {\sf GrB\_assign} ends and 
the dimension mismatch error listed above is returned.

From this point forward, in {\sf GrB\_NONBLOCKING} mode, the method can 
optionally exit with {\sf GrB\_SUCCESS} return code and defer any computation 
and/or execution error codes.

We are now ready to carry out the assign and any additional 
associated operations.  We describe this in terms of two intermediate vectors:
\begin{itemize}
    \item $\vector{\widetilde{t}}$: The vector holding the elements from
    $\vector{\widetilde{u}}$ in their destination locations relative to 
    $\vector{\widetilde{c}}$.
    
    \item $\vector{\widetilde{z}}$: The vector holding the result after 
    application of the (optional) accumulation.
\end{itemize}

The intermediate vector, $\vector{\widetilde{t}}$, is created as follows:
\[
\vector{\widetilde{t}} = \langle
\bold{D}({\sf u}), \bold{size}(\vector{\widetilde{c}}),
%\bold{L}(\vector{\widetilde{t}}) =
\{(\grbarray{\widetilde{J}}[j],\vector{\widetilde{u}}(j)) \ \forall \ j, \ 
0 \leq j < {\sf ncols} : j \in \bold{ind}(\vector{\widetilde{u}}) \} \rangle. 
\]
At this point, if any value of $\grbarray{\widetilde{J}}[j]$ is outside the valid 
range of indices for vector $\vector{\widetilde{c}}$, computation ends and the 
method returns the index out-of-bounds error listed above. In 
{\sf GrB\_NONBLOCKING} mode, the error can be deferred until a 
sequence-terminating {\sf GrB\_wait()} is called.  Regardless, the result 
matrix, {\sf C}, is invalid from this point forward in the 
sequence.

The intermediate vector $\vector{\widetilde{z}}$ is created as follows:
\begin{itemize}
    \item If ${\sf accum} = {\sf GrB\_NULL}$, then 
    $\vector{\widetilde{z}} = \vector{\widetilde{t}}$.

    \item If ${\sf accum} = \langle D_x, D_y, D_z, \odot \rangle$, then vector 
    $\vector{\widetilde{z}}$ is defined as 
        \[ 
        \vector{\widetilde{z}} =
        \langle D_z, \bold{size}(\vector{\widetilde{c}}), 
        %\bold{L}(\vector{\widetilde{z}}) = 
        \{(j,z_{j}) \ \forall \ j \in \bold{ind}(\vector{\widetilde{c}}) \cup 
        \bold{ind}(\vector{\widetilde{t}}) \} \rangle.
        \]
    The values of the elements of $\vector{\widetilde{z}}$ are computed based on the relationships between the sets of indices in $\vector{\widetilde{w}}$ and $\vector{\widetilde{t}}$.
\[
z_{j} = \vector{\widetilde{c}}(j) \odot \vector{\widetilde{t}}(j), \ \mbox{if}\  
j \in  (\bold{ind}(\vector{\widetilde{t}}) \cap \bold{ind}(\vector{\widetilde{c}})),
\]
\[
z_{j} = \vector{\widetilde{c}}(j), \ \mbox{if}\  j \in  (\bold{ind}(
\vector{\widetilde{c}}) - (\bold{ind}(\vector{\widetilde{t}}) \cap 
\bold{ind}(\vector{\widetilde{c}}))),
\]
\[
z_{j} = \vector{\widetilde{t}}(j), \ \mbox{if}\  j \in  (\bold{ind}(
\vector{\widetilde{t}}) - (\bold{ind}(\vector{\widetilde{t}}) \cap 
\bold{ind}(\vector{\widetilde{c}}))).
\]
where the difference operator in the previous expressions refers to set difference.
\end{itemize}

Finally, the set of output values that make up the $\vector{\widetilde{z}}$ 
vector are written into the column of the final result matrix, 
${\sf C}({\sf row\_index},:)$.  This is carried out under control of the mask 
which acts as a ``write mask''.
\begin{itemize}
    \item If {\sf desc[GrB\_OUTP].GrB\_REPLACE} is set then any values in 
    ${\sf C}({\sf row\_index},:)$ on input to {\sf GrB\_assign()} are deleted 
    and the new content of the column is given by:
    \[ 
		\bold{L}({\sf C}) = \{ (i,j,C_{ij}) : i \neq {\sf row\_index} \} \cup \{({\sf row\_index},j,z_{j}) : j \in 
    (\bold{ind}(\vector{\widetilde{z}}) \cap \bold{ind}(\vector{\widetilde{m}})) \}. 
    \]

    \item If {\sf desc[GrB\_OUTP].GrB\_REPLACE} is not set, the elements of 
    $\vector{\widetilde{z}}$ indicated by the mask are copied into the column 
    of the final result matrix, ${\sf C}({\sf row\_index},:)$, and elements of 
    this column that fall outside the set indicated by the mask are unchanged:
	\begin{eqnarray} 
		\bold{L}({\sf C}) & = & \{ (i,j,C_{ij}) : i \neq {\sf row\_index} \} \cup \nonumber \\
		& & \{({\sf row\_index},j,\vector{\widetilde{c}}(j)) : j \in (\bold{ind}(\vector{\widetilde{c}}) \cap \bold{ind}(\neg \vector{\widetilde{m}})) \} \cup \nonumber \\ 
		& & \{({\sf row\_index},j,z_{j})                     : j \in (\bold{ind}(\vector{\widetilde{z}}) \cap \bold{ind}(\vector{\widetilde{m}})) \}. \nonumber
	\end{eqnarray}
\end{itemize}

In {\sf GrB\_BLOCKING} mode, the method exits with return value 
{\sf GrB\_SUCCESS} and the new content of vector {\sf w} is as defined above
and fully computed.  
In {\sf GrB\_NONBLOCKING} mode, the method exits with return value 
{\sf GrB\_SUCCESS} and the new content of vector {\sf w} is as defined above 
but may not be fully computed; however, it can be used in the next GraphBLAS 
method call in a sequence.


%-----------------------------------------------------------------------------
\subsubsection{{\sf Matrix\_assign}: Constant matrix variant}

Assign the same value to a specified subgraph.  With the use of {\sf GrB\_ALL}, 
the entire destination matrix can be filled with the constant.

\paragraph{\syntax}

\begin{verbatim}
        GrB_Info GrB_Matrix_assign(GrB_Matrix              C,
                                   const GrB_Matrix        Mask,
                                   const GrB_BinaryOp      accum,
                                   <type>                  val,
                                   const GrB_Index        *row_indices,
                                   const GrB_Index         nrows,
                                   const GrB_Index        *col_indices,
                                   const GrB_Index         ncols,
                                   const GrB_Descriptor    desc);
\end{verbatim}

\paragraph{Parameters}

\begin{itemize}[leftmargin=1.1in]
    \item[{\sf C}]    ({\sf INOUT}) An existing GraphBLAS Matrix.  On input,
    the matrix provides values that may be accumulated with the result of the
    assign operation.  On output, this matrix holds the results of the
    operation.

    \item[{\sf Mask}] ({\sf IN}) An optional ``write'' mask that controls which
    results from this operation are stored into the output matrix {\sf C}. The 
    mask dimensions must match those of the matrix {\sf C} and the domain of the 
    {\sf Mask} matrix must be of type {\sf bool} or any of the predefined 
    ``built-in'' types in Table~\ref{Tab:PredefinedTypes}.  If the default
    matrix is desired (\ie, with correct dimensions and filled with {\sf true}), 
    {\sf GrB\_NULL} should be specified.

    \item[{\sf accum}]    ({\sf IN}) An optional operator used for accumulating
    entries into existing {\sf C} entries: ${\sf accum} = \langle D_x,
    D_y, D_z,\odot \rangle$. If assignment rather than accumulation is
    desired, {\sf GrB\_NULL} should be specified.

    \item[{\sf val}]    ({\sf IN}) Scalar value to assign to (a subset of) {\sf C}.
    
    \item[{\sf row\_indices}]  ({\sf IN}) Pointer to the ordered set (array) of 
    indices corresponding to the rows of {\sf C} that are assigned.  If all rows
    of {\sf C} are to be assigned in order from $0$ to ${\sf nrows} - 1$, then 
    {\sf GrB\_ALL} can be specified.  Regardless of execution mode and return 
    value, this array may be manipulated by the caller after this operation 
    returns without affecting any deferred computations for this operation.  
    Unlike other variants, if there are duplicated values in this array the 
    result is still defined.

    \item[{\sf nrows}] ({\sf IN}) The number of values in {\sf row\_indices} 
	array. Must be in the range: $[0, \bold{nrows}({\sf C})]$.  If
    {\sf nrows} is zero, the operation becomes a NO-OP.

    \item[{\sf col\_indices}]  ({\sf IN}) Pointer to the ordered set (array) of 
    indices corresponding to the columns of {\sf C} that are assigned.  If all columns
    of {\sf C} are to be assigned in order from $0$ to ${\sf ncols} - 1$, then 
    {\sf GrB\_ALL} can be specified.  Regardless of execution mode and return 
    value, this array may be manipulated by the caller after this operation 
    returns without affecting any deferred computations for this operation.
    Unlike other variants, if there are duplicated values in this array the 
    result is still defined.

    \item[{\sf ncols}] ({\sf IN}) The number of values in {\sf col\_indices} 
	array. Must be in the range: $[0, \bold{ncols}({\sf C})]$.  If
    {\sf ncols} is zero, the operation becomes a NO-OP.

    \item[{\sf desc}] ({\sf IN}) An optional operation descriptor. If
    a \emph{default} descriptor is desired, {\sf GrB\_NULL} should be
    specified. Non-default field/value pairs are listed as follows:  \\

    \begin{tabular}{lllp{2.5in}}
    Param & Field  & Value & Description \\
    \hline
        {\sf C}    & {\sf GrB\_OUTP} & {\sf GrB\_REPLACE} & Output matrix {\sf C} 
        is cleared (all elements removed) before result is stored in it.\\
    
        {\sf mask} & {\sf GrB\_MASK} & {\sf GrB\_SCMP} & Use the structural 
        complement of {\sf Mask}. \\
    \end{tabular}

\end{itemize}

\paragraph{Return Values}

\begin{itemize}[leftmargin=2.1in]
    \item[{\sf GrB\_SUCCESS}]         In blocking mode, the operation completed
    successfully. In non-blocking mode, this indicates that the compatibility 
    tests on dimensions and domains for the input arguments passed successfully. 
    Either way, output matrix {\sf C} is ready to be used in the next method of 
    the sequence.

    \item[{\sf GrB\_PANIC}]            Unknown internal error.
    
    \item[{\sf GrB\_INVALID\_OBJECT}] This is returned in any execution mode 
    whenever one of the opaque GraphBLAS objects (input or output) is in an invalid 
    state caused by a previous execution error.  Call {\sf GrB\_error()} to access 
    any error messages generated by the implementation.

    \item[{\sf GrB\_OUT\_OF\_MEMORY}]  Not enough memory available for operation.

    \item[{\sf GrB\_UNINITIALIZED\_OBJECT}] One or more of the GraphBLAS objects
    has not been initialized by a call to {\sf new} (or {\sf dup} for vector
    parameters).

    \item[{\sf GrB\_INDEX\_OUT\_OF\_BOUNDS}]  A value in {\sf row\_indices} is greater
    than or equal to $\bold{nrows}({\sf C})$, or a value in {\sf col\_indices} is greater
    than or equal to $\bold{ncols}({\sf C})$.  In non-blocking mode, this can be
    reported as an execution error.

    \item[{\sf GrB\_DIMENSION\_MISMATCH}]  {\sf Mask} and {\sf C} dimensions are
    incompatible, {\sf nrows} is not less than $\bold{nrows}({\sf C})$, or
    {\sf ncols} is not less than $\bold{ncols}({\sf C})$. 

    \item[{\sf GrB\_DOMAIN\_MISMATCH}]     The domains of the matrix and scalar are
	incompatible with each other or the corresponding domains of the 
    accumulation operator, or the mask's domain is not compatible with {\sf bool}.

    \item[{\sf GrB\_NULL\_POINTER}] Either argument {\sf row\_indices} is a {\sf NULL} pointer,
	    argument {\sf col\_indices} is a {\sf NULL} pointer, or both.
\end{itemize}


\paragraph{Description}


This variant of {\sf GrB\_assign} computes the result of assigning a constant
scalar value to locations in a destination GraphBLAS matrix: 
${\sf C}({\sf row\_indices, col\_indices}) = {\sf val}$; or, if an optional 
binary accumulation operator ($\odot$) is provided, 
${\sf C}({\sf row\_indices, col\_indices}) = 
{\sf w}({\sf row\_indices, col\_indices}) \odot {\sf val}$.  
More explicitly:
\[
\begin{aligned}
	{\sf C}({\sf row\_indices}[i], {\sf col\_indices}[j]) =\ & {\sf val} \mbox{~~or~~} \\
    {\sf C}({\sf row\_indices}[i], {\sf col\_indices}[j]) =\ & 
              {\sf C}({\sf row\_indices}[i], {\sf col\_indices}[j]) \odot {\sf val} \\
    & \forall \ (i,j) \ : \ 0 \leq i < {\sf nrows},  0 \leq j < {\sf ncols}
\end{aligned}
\]  
Logically, this operation occurs in three steps:
\begin{enumerate}[leftmargin=0.75in]
\item[Setup] The internal vectors and mask used in the computation are formed 
and their domains and dimensions are tested for compatibility.
\item[Compute] The indicated computations are carried out.
\item[Output] The result is written into the output matrix, possibly under 
control of a mask.
\end{enumerate}

Up to two argument matrices are used in the {\sf GrB\_assign} operation:
\begin{enumerate}
	\item ${\sf C} = \langle \bold{D}({\sf C}),\bold{nrows}({\sf C}),
    \bold{ncols}({\sf C}),\bold{L}({\sf C}) = \{(i,j,C_{ij}) \} \rangle$
    
	\item ${\sf Mask} = \langle \bold{D}({\sf Mask}),\bold{nrows}({\sf Mask}),
    \bold{ncols}({\sf Mask}),\bold{L}({\sf Mask}) = \{(i,j,M_{ij}) \} \rangle$ (optional)
\end{enumerate}

The argument scalar, matrices, and the accumulation 
operator (if provided) are tested for domain compatibility as follows:
\begin{enumerate}
	\item The domain of {\sf Mask} (if not {\sf GrB\_NULL}) must be from one of 
    the pre-defined types of Table~\ref{Tab:PredefinedTypes}.

	\item If {\sf accum} is {\sf GrB\_NULL}, then $\bold{D}({\sf C})$ must be 
    compatible with $\bold{D}({\sf val})$.

	\item If {\sf accum} is not {\sf GrB\_NULL}, then $\bold{D}({\sf C})$ must be
    compatible with $D_x$ and $D_z$ of the accumulation operator and 
    $\bold{D}({\sf val})$ must be compatible with $D_y$ of the accumulation operator.
\end{enumerate}
Two domains are compatible with each other if values from one domain can be cast 
to values in the other domain as per the rules of the C language.
In particular, domains from Table~\ref{Tab:PredefinedTypes} are all compatible 
with each other. A domain from a user-defined type is only compatible with itself.
If any compatibility rule above is violated, execution of {\sf GrB\_assign} ends
and the domain mismatch error listed above is returned.

From the arguments, the internal matrices, index arrays, and mask used in 
the computation are formed ($\leftarrow$ denotes copy):
\begin{enumerate}
	\item Matrix $\matrix{\widetilde{C}} \leftarrow {\sf C}$.

	\item Two-dimensional mask $\matrix{\widetilde{M}}$ is computed from 
    argument {\sf Mask} as follows:
	\begin{enumerate}

		\item	If ${\sf Mask} = {\sf GrB\_NULL}$, then $\matrix{\widetilde{M}} = 
        \langle \bold{nrows}({\sf C}), \bold{ncols}({\sf C}), \{(i,j), 
        \forall i,j : 0 \leq i <  \bold{nrows}({\sf C}), 0 \leq j < 
        \bold{ncols}({\sf C}) \} \rangle$.

		\item	Otherwise, $\matrix{\widetilde{M}} = \langle 
        \bold{nrows}({\sf Mask}), \bold{ncols}({\sf Mask}), \{(i,j) : 
        (i,j) \in \bold{ind}({\sf Mask}) \wedge 
        ({\sf bool}){\sf Mask}(i,j) = \true\} \rangle$.

		\item	If ${\sf desc[GrB\_MASK].GrB\_SCMP}$ is set, then 
        $\matrix{\widetilde{M}} \leftarrow \neg \matrix{\widetilde{M}}$.
	\end{enumerate}

    \item The internal row index array, $\grbarray{\widetilde{I}}$, is computed from 
    argument {\sf row\_indices} as follows:
	\begin{enumerate}
		\item	If ${\sf row\_indices} = {\sf GrB\_ALL}$, then 
        $\grbarray{\widetilde{I}}[i] = i, \forall i : 0 \leq i < {\sf nrows}$.

		\item	Otherwise, $\grbarray{\widetilde{I}}[i] = {\sf row\_indices}[i], 
        \forall i : 0 \leq i < {\sf nrows}$.
    \end{enumerate}
    
    \item The internal column index array, $\grbarray{\widetilde{J}}$, is computed from 
    argument {\sf col\_indices} as follows:
	\begin{enumerate}
		\item	If ${\sf col\_indices} = {\sf GrB\_ALL}$, then 
        $\grbarray{\widetilde{J}}[j] = j, \forall j : 0 \leq j < {\sf ncols}$.

		\item	Otherwise, $\grbarray{\widetilde{J}}[j] = {\sf col\_indices}[j], 
        \forall j : 0 \leq j < {\sf ncols}$.
    \end{enumerate}
\end{enumerate}

The internal matrix and mask are checked for dimension compatibility. 
The following conditions must hold:
\begin{enumerate}
    \item $\bold{nrows}(\matrix{\widetilde{C}}) = \bold{nrows}(\matrix{\widetilde{M}})$.

    \item $\bold{ncols}(\matrix{\widetilde{C}}) = \bold{ncols}(\matrix{\widetilde{M}})$.

    \item $0 \leq {\sf nrows} \leq \bold{nrows}(\matrix{\widetilde{C}})$.

    \item $0 \leq {\sf ncols} \leq \bold{ncols}(\matrix{\widetilde{C}})$.
\end{enumerate}
If any compatibility rule above is violated, execution of {\sf GrB\_assign} ends and 
the dimension mismatch error listed above is returned.

From this point forward, in {\sf GrB\_NONBLOCKING} mode, the method can 
optionally exit with {\sf GrB\_SUCCESS} return code and defer any computation 
and/or execution error codes.

We are now ready to carry out the assign and any additional 
associated operations.  We describe this in terms of two intermediate vectors:
\begin{itemize}
	\item $\matrix{\widetilde{T}}$: The matrix holding the copies of the scalar 
    {\sf val} in their destination locations relative to 
    $\matrix{\widetilde{C}}$.
    
	\item $\matrix{\widetilde{Z}}$: The matrix holding the result after 
    application of the (optional) accumulation operator.
\end{itemize}

The intermediate matrix, $\matrix{\widetilde{T}}$, is created as follows:
\[
\begin{aligned}
\matrix{\widetilde{T}} = \langle & \bold{D}({\sf val}),
                           \bold{nrows}(\matrix{\widetilde{C}}), 
                           \bold{ncols}(\matrix{\widetilde{C}}), \\
 & \{ (\grbarray{\widetilde{I}}[i],\grbarray{\widetilde{J}}[j], {\sf val}) 
 \ \forall \ (i,j), \ 0 \leq i < {\sf nrows}, \ 0 \leq j < {\sf ncols} \} \rangle. 
\end{aligned}
\]
If either $\grbarray{\widetilde{I}}$ or $\grbarray{\widetilde{J}}$ is empty, this 
operation results in an empty matrix, $\matrix{\widetilde{T}}$.  Otherwise, if 
any value in the $\grbarray{\widetilde{I}}$ array is not in
the range $[0,\ \bold{nrows}(\matrix{\widetilde{C}}) )$ or any value in the 
$\grbarray{\widetilde{J}}$ array is not in the range 
$[0,\ \bold{ncols}(\matrix{\widetilde{C}}))$, the execution of {\sf GrB\_assign} 
ends and the index out-of-bounds error listed above is generated. In 
{\sf GrB\_NONBLOCKING} mode, the error can be deferred until a 
sequence-terminating {\sf GrB\_wait()} is called.  Regardless, the result 
matrix {\sf C} is invalid from this point forward in the sequence.

The intermediate matrix $\matrix{\widetilde{Z}}$ is created as follows:
\begin{itemize}
    \item If ${\sf accum} = {\sf GrB\_NULL}$, then 
    $\matrix{\widetilde{Z}} = \matrix{\widetilde{T}}$.

    \item If ${\sf accum} = \langle D_x, D_y, D_z, \odot \rangle$, then matrix 
    $\matrix{\widetilde{Z}}$ is defined as 
        \[ 
        \matrix{\widetilde{Z}} = 
        \langle D_z, \bold{nrows}(\matrix{\widetilde{C}}),
        \bold{ncols}(\matrix{\widetilde{C}}), 
        %\bold{L}(\matrix{\widetilde{Z}}) =
	    \{(i,j,Z_{ij}) \ \forall \ (i,j) \in \bold{ind}(\matrix{\widetilde{C}}) \cup 
        \bold{ind}(\matrix{\widetilde{T}}) \} \rangle.\]
    The values of the elements of $\matrix{\widetilde{Z}}$ are computed based on 
    the relationships between the sets of indices in $\matrix{\widetilde{C}}$ and 
    $\matrix{\widetilde{T}}$.
\[
	Z_{ij} = \matrix{\widetilde{C}}(i,j) \odot \matrix{\widetilde{T}}(i,j), 
    \mbox{~if~} (i,j) \in  (\bold{ind}(\matrix{\widetilde{T}}) \cap 
    \bold{ind}(\matrix{\widetilde{C}})),
\]
\[
	Z_{ij} = \matrix{\widetilde{C}}(i,j), \mbox{~if~} (i,j) \in  
    (\bold{ind}(\matrix{\widetilde{C}}) - (\bold{ind}(\matrix{\widetilde{T}}) \cap 
    \bold{ind}(\matrix{\widetilde{C}}))),
\]
\[
	Z_{ij} = \matrix{\widetilde{T}}(i,j), \mbox{~if~} (i,j) \in  
    (\bold{ind}(\matrix{\widetilde{T}}) - (\bold{ind}(\matrix{\widetilde{T}}) \cap 
    \bold{ind}(\matrix{\widetilde{C}}))).
\]
where the difference operator in the previous expressions refers to set difference.
\end{itemize}

Finally, the set of output values that make up the $\matrix{\widetilde{Z}}$ 
matrix are written into the final result matrix, ${\sf C}$. 
This is carried out under control of the mask which acts as a ``write mask''.
\begin{itemize}
\item If {\sf desc[GrB\_OUTP].GrB\_REPLACE} is set then any values in ${\sf C}$ 
on input to {\sf GrB\_assign()} are deleted and the new output matrix ${\sf C}$ is,
		\[ \bold{L}({\sf C}) = \{(i,j,Z_{ij}) : (i,j) \in (\bold{ind}(\matrix{\widetilde{Z}}) 
\cap \bold{ind}(\matrix{\widetilde{M}})) \}. \]

\item If {\sf desc[GrB\_OUTP].GrB\_REPLACE} is not set, the elements of 
$\matrix{\widetilde{Z}}$ indicated by 
the mask are copied into the result matrix, ${\sf C}$, and elements of 
${\sf C}$ that fall outside the set indicated by the mask are unchanged:
		\[ \bold{L}({\sf C}) = \{(i,j,C_{ij}) : (i,j) \in (\bold{ind}(\matrix{\sf C}) 
		\cap \bold{ind}(\neg \matrix{\widetilde{M}})) \} \cup \{(i,j,Z_{ij}) : (i,j) \in 
(\bold{ind}(\matrix{\widetilde{Z}}) \cap \bold{ind}(\matrix{\widetilde{M}})) \}. \]
\end{itemize}

In {\sf GrB\_BLOCKING} mode, the method exits with return value 
{\sf GrB\_SUCCESS} and the new content of matrix {\sf C} is as defined above
and fully computed.  
In {\sf GrB\_NONBLOCKING} mode, the method exits with return value 
{\sf GrB\_SUCCESS} and the new content of matrix {\sf C} is as defined above 
but may not be fully computed; however, it can be used in the next GraphBLAS 
method call in a sequence.
