%=========================================================================

\subsection{{\sf reduce}: Perform a reduction across the elements of an object}

Computes the reduction of the values of the elements of a vector or matrix.

%-----------------------------------------------------------------------------

\subsubsection{{\sf reduce}: Standard matrix to vector variant}

This performs a reduction across rows of a matrix to produce a vector.  If 
reduction down columns is desired, the input matrix should be transposed 
using the descriptor.

\paragraph{\syntax}

\begin{verbatim}
        GrB_Info GrB_reduce(GrB_Vector             w,
                            const GrB_Vector       mask,
                            const GrB_BinaryOp     accum,
                            const GrB_Monoid       op,  
                            const GrB_Matrix       A,
                            const GrB_Descriptor   desc);
                            
        GrB_Info GrB_reduce(GrB_Vector             w,
                            const GrB_Vector       mask,
                            const GrB_BinaryOp     accum,
                            const GrB_BinaryOp     op,  
                            const GrB_Matrix       A,
                            const GrB_Descriptor   desc);
\end{verbatim}

\paragraph{Parameters}

\begin{itemize}[leftmargin=1.1in]
    \item[{\sf w}]    ({\sf INOUT}) An existing GraphBLAS vector.  On input,
    the vector provides values that may be accumulated with the result of the
    reduction operation.  On output, this vector holds the results of the
    operation.

    \item[{\sf mask}] ({\sf IN}) An optional ``write'' mask that controls which
    results from this operation are stored into the output vector {\sf w}. The 
    mask dimensions must match those of the vector {\sf w}. If the 
    {\sf GrB\_STRUCTURE} descriptor is {\em not} set for the mask, the domain of the
    {\sf mask} vector must be of type {\sf bool} or any of the predefined 
    ``built-in'' types in Table~\ref{Tab:PredefinedTypes}.  If the default
    mask is desired (\ie, a mask that is all {\sf true} with the dimensions of {\sf w}), 
    {\sf GrB\_NULL} should be specified.

    \item[{\sf accum}] ({\sf IN}) An optional binary operator used for accumulating
    entries into existing {\sf w} entries. If assignment rather than accumulation is
    desired, {\sf GrB\_NULL} should be specified.

    \item[{\sf op}]    ({\sf IN}) The monoid or binary operator 
    used in the element-wise reduction operation.  Depending on which type is
    passed, the following defines the binary operator with one domain, $F_b=\langle D,D,D,\oplus \rangle$, that is used:
    \begin{itemize}[leftmargin=1.1in]
    \item[BinaryOp:] $F_b = \langle \bDout({\sf op}), \bDin1({\sf op}),
    \bDin2({\sf op}),\bold{\bigodot}({\sf op})\rangle$.  
    \item[Monoid:] $F_b = \langle \bold{D}({\sf op}), \bold{D}({\sf op}),
    \bold{D}({\sf op}),\bold{\bigodot}({\sf op})\rangle$,
    the identity element of the monoid is ignored. 
    \end{itemize}
    If {\sf op} is a {\sf GrB\_BinaryOp}, then all its domains must be the same.
    Furthermore, in both cases $\bold{\bigodot}({\sf op})$ must be commutative and associative. Otherwise, the outcome of the operation is undefined.
    
    \item[{\sf A}]     ({\sf IN}) The GraphBLAS matrix on which
	    reduction will be performed.

    \item[{\sf desc}] ({\sf IN}) An optional operation descriptor. If
    a \emph{default} descriptor is desired, {\sf GrB\_NULL} should be
    specified. Non-default field/value pairs are listed as follows:  \\

    \hspace*{-2em}\begin{tabular}{lllp{2.7in}}
        Param & Field  & Value & Description \\
        \hline
        {\sf w}    & {\sf GrB\_OUTP} & {\sf GrB\_REPLACE} & Output vector {\sf w}
        is cleared (all elements removed) before the result is stored in it.\\

        {\sf mask} & {\sf GrB\_MASK} & {\sf GrB\_STRUCTURE}   & The write mask is
        constructed from the structure (pattern of stored values) of the input
        {\sf mask} vector. The stored values are not examined.\\

        {\sf mask} & {\sf GrB\_MASK} & {\sf GrB\_COMP}   & Use the 
        complement of {\sf mask}. \\

        {\sf A}    & {\sf GrB\_INP0} & {\sf GrB\_TRAN}   & Use transpose of {\sf A}
        for the operation. \\
    \end{tabular}
\end{itemize}

\paragraph{Return Values}

\begin{itemize}[leftmargin=2.1in]
    \item[{\sf GrB\_SUCCESS}]         In blocking mode, the operation completed
    successfully. In non-blocking mode, this indicates that the compatibility 
    tests on dimensions and domains for the input arguments passed successfully. 
    Either way, output vector {\sf w} is ready to be used in the next method of 
    the sequence.

    \item[{\sf GrB\_PANIC}]           Unknown internal error.

    \item[{\sf GrB\_INVALID\_OBJECT}] This is returned in any execution mode 
    whenever one of the opaque GraphBLAS objects (input or output) is in an invalid 
    state caused by a previous execution error.  Call {\sf GrB\_error()} to access 
    any error messages generated by the implementation.

    \item[{\sf GrB\_OUT\_OF\_MEMORY}] Not enough memory available for the operation.

    \item[{\sf GrB\_UNINITIALIZED\_OBJECT}] One or more of the GraphBLAS objects
    has not been initialized by a call to {\sf new} (or {\sf dup} for vector
    parameters).
    
    \item[{\sf GrB\_DIMENSION\_MISMATCH}]  {\sf mask}, {\sf w} and/or {\sf u} dimensions are
    incompatible. 

    \item[{\sf GrB\_DOMAIN\_MISMATCH}]    Either the domains of the various vectors and matrices are
    incompatible with the corresponding domains of the accumulation operator or 
    reduce function, or the domains of the GraphBLAS binary operator {\sf op} are 
    not all the same, or the mask's domain is not compatible with {\sf bool}
    (in the case where {\sf desc[GrB\_MASK].GrB\_STRUCTURE} is not set).
\end{itemize}

\paragraph{Description}

This variant of {\sf GrB\_reduce} computes the result of performing a reduction 
across each of the rows of an input matrix: ${\sf w}(i) = \bigoplus {\sf A}(i,:) \forall i$; 
or, if an optional binary accumulation operator is provided, 
${\sf w}(i) = {\sf w}(i) \odot (\bigoplus {\sf A}(i,:)) \forall i$, 
where $\bigoplus = \bigodot({F_b})$ and $\odot = \bigodot({\sf accum})$.

Logically, this operation occurs in three steps:
\begin{enumerate}[leftmargin=0.75in]
\item[\bf Setup] The internal vector, matrix and mask used in the computation are formed 
and their domains and dimensions are tested for compatibility.
\item[\bf Compute] The indicated computations are carried out.
\item[\bf Output] The result is written into the output vector, possibly under 
control of a mask.
\end{enumerate}

Up to two vector and one matrix argument are used in this {\sf GrB\_reduce} operation:
\begin{enumerate}
	\item ${\sf w} = \langle \bold{D}({\sf w}),\bold{size}({\sf w}),
    \bold{L}({\sf w}) = \{(i,w_i) \} \rangle$

	\item ${\sf mask} = \langle \bold{D}({\sf mask}),\bold{size}({\sf mask}),
    \bold{L}({\sf mask}) = \{(i,m_i) \} \rangle$ (optional)

	\item ${\sf A} = \langle \bold{D}({\sf A}),\bold{nrows}({\sf A}),
    \bold{ncols}({\sf A}),\bold{L}({\sf A}) = \{(i,j,A_{ij}) \} \rangle$
\end{enumerate}

The argument vector, matrix, reduction operator and accumulation 
operator (if provided) are tested for domain compatibility as follows:
\begin{enumerate}
	\item If {\sf mask} is not {\sf GrB\_NULL}, and ${\sf desc[GrB\_MASK].GrB\_STRUCTURE}$
    is not set, then $\bold{D}({\sf mask})$ must be from one of the pre-defined types of 
    Table~\ref{Tab:PredefinedTypes}.

	\item $\bold{D}({\sf w})$ must be 
    compatible with the domain of the reduction binary operator, $\bold{D}(F_b)$.

	\item If {\sf accum} is not {\sf GrB\_NULL}, then $\bold{D}({\sf w})$ must be
    compatible with $\bDin1({\sf accum})$ and $\bDout({\sf accum})$ of the accumulation operator and 
    $\bold{D}(F_b)$, must be compatible with $\bDin2({\sf accum})$ of the accumulation operator.

	\item $\bold{D}({\sf A})$ must be compatible with the domain of the binary reduction operator, $\bold{D}(F_b)$.
\end{enumerate}
Two domains are compatible with each other if values from one domain can be cast 
to values in the other domain as per the rules of the C language.
In particular, domains from Table~\ref{Tab:PredefinedTypes} are all compatible 
with each other. A domain from a user-defined type is only compatible with itself.
If any compatibility rule above is violated, execution of {\sf GrB\_reduce} ends
and the domain mismatch error listed above is returned.

From the argument vectors, the internal vectors and mask used in 
the computation are formed ($\leftarrow$ denotes copy):
\begin{enumerate}
	\item Vector $\vector{\widetilde{w}} \leftarrow {\sf w}$.

	\item One-dimensional mask, $\vector{\widetilde{m}}$, is computed from 
    argument {\sf mask} as follows:
	\begin{enumerate}
		\item If ${\sf mask} = {\sf GrB\_NULL}$, then $\vector{\widetilde{m}} = 
        \langle \bold{size}({\sf w}), \{i, \ \forall \ i : 0 \leq i < 
        \bold{size}({\sf w}) \} \rangle$.

		\item If {\sf mask} $\ne$ {\sf GrB\_NULL},  
        \begin{enumerate}
            \item If ${\sf desc[GrB\_MASK].GrB\_STRUCTURE}$ is set, then
            $\vector{\widetilde{m}} = 
            \langle \bold{size}({\sf mask}), \{i : i \in \bold{ind}({\sf mask}) \} \rangle$,
            \item Otherwise, $\vector{\widetilde{m}} = 
            \langle \bold{size}({\sf mask}), \{i : i \in \bold{ind}({\sf mask}) \wedge
            ({\sf bool}){\sf mask}(i) = \true \} \rangle$.
        \end{enumerate}

		\item	If ${\sf desc[GrB\_MASK].GrB\_COMP}$ is set, then 
        $\vector{\widetilde{m}} \leftarrow \neg \vector{\widetilde{m}}$.
	\end{enumerate}

	\item Matrix $\matrix{\widetilde{A}} \leftarrow 
    {\sf desc[GrB\_INP0].GrB\_TRAN} \ ? \ {\sf A}^T : {\sf A}$.
\end{enumerate}

The internal vectors and masks are checked for dimension compatibility. 
The following conditions must hold:
\begin{enumerate}
	\item $\bold{size}(\vector{\widetilde{w}}) = \bold{size}(\vector{\widetilde{m}})$
    \item $\bold{size}(\vector{\widetilde{w}}) = \bold{nrows}(\matrix{\widetilde{A}})$.
\end{enumerate}
If any compatibility rule above is violated, execution of {\sf GrB\_reduce} ends and 
the dimension mismatch error listed above is returned.

From this point forward, in {\sf GrB\_NONBLOCKING} mode, the method can 
optionally exit with {\sf GrB\_SUCCESS} return code and defer any computation 
and/or execution error codes.

We carry out the reduce and any additional 
associated operations.  We describe this in terms of two intermediate vectors:
\begin{itemize}
    \item $\vector{\widetilde{t}}$: The vector holding the result from reducing along the rows of input matrix
    $\matrix{\widetilde{A}}$.

    \item $\vector{\widetilde{z}}$: The vector holding the result after 
    application of the (optional) accumulation operator.
\end{itemize}

The intermediate vector, $\vector{\widetilde{t}}$, is created as follows:
\[
\vector{\widetilde{t}} = \langle
\bold{D}({\sf op}), \bold{size}(\vector{\widetilde{w}}),
\{(i,t_i) : \bold{ind}(A(i,:)) \neq \emptyset   \} \rangle. 
\]
The value of each of its elements is computed by
\[
	t_i = \bigoplus_{j \in \bold{ind}(\matrix{\widetilde{A}}(i,:))} \matrix{\widetilde{A}}(i,j), 
\]
where $\bigoplus = \bigodot(F_b)$.

\input{ops_accum_z_vector}

\input{ops_mask_replace_vector}


%-----------------------------------------------------------------------------

\subsubsection{{\sf reduce}: Vector-scalar variant}
\label{Sec:Reduce_vector_scalar}

Reduce all stored values into a single scalar.

\paragraph{\syntax}

\begin{verbatim}
        // scalar value + monoid (only)
        GrB_Info GrB_reduce(<type>               *val,
                            const GrB_BinaryOp    accum,
                            const GrB_Monoid      op,
                            const GrB_Vector      u,
                            const GrB_Descriptor  desc);

        // GraphBLAS Scalar + monoid
        GrB_Info GrB_reduce(GrB_Scalar            s,
                            const GrB_BinaryOp    accum,
                            const GrB_Monoid      op,
                            const GrB_Vector      u,
                            const GrB_Descriptor  desc);

        // GraphBLAS Scalar + binary operator
        GrB_Info GrB_reduce(GrB_Scalar            s,
                            const GrB_BinaryOp    accum,
                            const GrB_BinaryOp    op,
                            const GrB_Vector      u,
                            const GrB_Descriptor  desc);
\end{verbatim}

\paragraph{Parameters}

\begin{itemize}[leftmargin=1.1in]
    \item[{\sf val} or {\sf s}] ({\sf INOUT}) Scalar to store final reduced value into. On input,
    the scalar provides a value that may be accumulated (optionally) with the result of the
    reduction operation.  On output, this scalar holds the results of the
    operation.

    \item[{\sf accum}] ({\sf IN}) An optional binary operator used for accumulating
    entries into an existing scalar ({\sf s} or {\sf val}) value. If assignment rather than accumulation is
    desired, {\sf GrB\_NULL} should be specified.

    \item[{\sf op}]    ({\sf IN}) The monoid ($M = \langle D,\oplus,0 \rangle$) or binary operator 
    ($F_b = \langle D, D, D, \oplus \rangle$) used in the reduction operation. The $\oplus$ operator 
    must be commutative and associative; otherwise, the outcome of the operation is undefined.  
    %\scott{Note the restrictions on the domains of the binary operator
    %$F_b = \langle D, D, D, \oplus \rangle$, plus lots comments from the  implementors
    %about the use of binary ops/lack of identity leading to poor performance.}
    
    \item[{\sf u}]     ({\sf IN}) The GraphBLAS vector on which
    reduction will be performed.

    \item[{\sf desc}] ({\sf IN}) An optional operation descriptor. If
    a \emph{default} descriptor is desired, {\sf GrB\_NULL} should be
    specified. Non-default field/value pairs are listed as follows:  \\

    \begin{tabular}{lllp{2.5in}}
        Param & Field  & Value & Description \\
        \hline
    \end{tabular}

    \emph{Note:} This argument is defined for consistency with the other GraphBLAS operations.
    There are currently no non-default field/value pairs that can be set for this operation.
\end{itemize}

\paragraph{Return Values}

\begin{itemize}[leftmargin=2.1in]
    \item[{\sf GrB\_SUCCESS}]         In blocking or non-blocking mode, 
    the operation completed successfully, and the
    output scalar ({\sf s} or {\sf val}) is ready to be used in the next method of 
    the sequence.

    \item[{\sf GrB\_PANIC}]           Unknown internal error.

    \item[{\sf GrB\_INVALID\_OBJECT}] This is returned in any execution mode 
    whenever one of the opaque GraphBLAS objects (input or output) is in an invalid 
    state caused by a previous execution error.  Call {\sf GrB\_error()} to access 
    any error messages generated by the implementation.

    \item[{\sf GrB\_OUT\_OF\_MEMORY}] Not enough memory available for the operation.

    \item[{\sf GrB\_UNINITIALIZED\_OBJECT}] One or more of the GraphBLAS objects 
    has not been initialized by a call to a respective constructor.

    \item[{\sf GrB\_NULL\_POINTER}]  {\sf val} pointer is {\sf NULL}.
    
    \item[{\sf GrB\_DOMAIN\_MISMATCH}]    The domains of input and output arguments are
    incompatible with the corresponding domains of the accumulation operator, 
    or reduce operator.
\end{itemize}

\paragraph{Description}

This variant of {\sf GrB\_reduce} computes the result of performing a reduction
across all of the stored elements of an input vector storing the result into either
{\sf s} or {\sf val}.  This corresponds to (shown here for the scalar value case only):
\begin{equation*}
{\sf val} ~ = 
  \begin{cases}
     \hspace{3.7em} \bigoplus_{i \in \mathbf{ind}({\sf u})} {\sf u}(i) \text{,~~~ or} \\& \\
    {\sf val} ~\odot~ \left[ \bigoplus_{i \in \mathbf{ind}({\sf u})} {\sf u}(i)\right], \text{~if the the optional accumulator is specified.} 
  \end{cases}
\end{equation*}
where $\bigoplus = \bigodot({\sf op})$ and $\odot = \bigodot({\sf accum})$.  

Logically, this operation occurs in three steps:
\begin{enumerate}[leftmargin=0.85in]
\item[\bf Setup] The internal vector used in the computation is formed 
and its domain is tested for compatibility.
\item[\bf Compute] The indicated computations are carried out.
\item[\bf Output] The result is written into the output scalar.
\end{enumerate}

One vector argument is used in this {\sf GrB\_reduce} operation:
\begin{enumerate}
	\item ${\sf u} = \langle \bold{D}({\sf u}),\bold{size}({\sf u}),
    \bold{L}({\sf u}) = \{(i,u_i) \} \rangle$
\end{enumerate}

The output scalar, argument vector, reduction operator and accumulation 
operator (if provided) are tested for domain compatibility as follows:
\begin{enumerate}

	\item If {\sf accum} is {\sf GrB\_NULL}, then $\bold{D}({\sf val})$ or $\bold{D}({\sf s})$ must be
    compatible with $\bold{D}({\sf op})$ from $M$  (or with $\bDin1({\sf op})$ and 
    $\bDin2({\sf op})$ from $F_b$).

	\item If {\sf accum} is not {\sf GrB\_NULL}, then $\bold{D}({\sf val})$ or $\bold{D}({\sf s})$ must be
    compatible with $\bDin1({\sf accum})$ and $\bDout({\sf accum})$ of the accumulation operator, and 
    $\bold{D}({\sf op})$ from $M$  (or $\bDout({\sf op})$ from $F_b$) must be compatible 
    with $\bDin2({\sf accum})$ of the accumulation operator.

	\item $\bold{D}({\sf u})$ must be compatible with $\bold{D}({\sf op})$ from $M$ 
    (or with $\bDin1({\sf op})$ and $\bDin2({\sf op})$ from $F_b$).
\end{enumerate}
Two domains are compatible with each other if values from one domain can be cast 
to values in the other domain as per the rules of the C language.
In particular, domains from Table~\ref{Tab:PredefinedTypes} are all compatible 
with each other. A domain from a user-defined type is only compatible with itself.
If any compatibility rule above is violated, execution of {\sf GrB\_reduce} ends and 
the domain mismatch error listed above is returned.

The number of values stored in the input, {\sf u}, is checked.  If there are no stored values in {\sf u}, then one of the following occurs depending on the output variant:
\[ 
{\sf {\bf L}(s)} =
\begin{cases}
    \{\}, & \mbox{(cleared) if {\sf accum} = {\sf GrB\_NULL}}, ~~\\ & \\
    {\sf {\bf L}(s)}, ~~~ ~~~~~~~ & \mbox{(unchanged) otherwise,}
\end{cases}
\]
or
\[ 
{\sf val} =
\begin{cases}
    \bold{0}({\sf op}), & \mbox{(cleared) if {\sf accum} = {\sf GrB\_NULL}}, \\ & \\
    {\sf val}~\odot~ {\bf 0}({\sf op}), & \mbox{otherwise,}
\end{cases}
\]
where $\bold{0}({\sf op})$ is the identity of the monoid. The operation returns immediately with {\sf GrB\_SUCCESS}.

For all other cases, the internal vector and scalar used in 
the computation is formed ($\leftarrow$ denotes copy):
\begin{enumerate}
	\item Vector $\vector{\widetilde{u}} \leftarrow {\sf u}$.
    \item Scalar $\scalar{\tilde{s}} \leftarrow {\sf s}$ (GraphBLAS scalar case).
\end{enumerate}
% Note that these copies are not really necessary as aliasing cannot really happen 
% but this section is here to be consistent with the format of other operations.

We are now ready to carry out the reduction and any additional associated operations.  
An intermediate scalar result $t$ is computed as follows:
\[ 
t ~ = ~
    \bigoplus\limits_{i \in \bold{ind}(\vector{\widetilde{u}})} \vector{\widetilde{u}}(i),
\]
where $\oplus = \bigodot({\sf op})$.

The final reduction value is computed as follows:
\[
{\sf {\bf L}(s)} \leftarrow
    \begin{cases}
    \{ t \}, & \text{ when {\sf accum = GrB\_NULL} or $\scalar{\tilde{s}}$ is empty, or} \\ & \\
    \{ {\bf val}(\scalar{\tilde{s}}) ~\odot ~ t \}, & \text{otherwise;} \\
    \end{cases}
\]
or 
\[
{\sf val} \leftarrow
    \begin{cases}
    t, & \text{ when {\sf accum = GrB\_NULL}, or} \hspace*{0.95in} \\ & \\
    {\sf val} ~\odot ~t, ~~~~~ & \text{otherwise;} \\
    \end{cases}
\]

In both {\sf GrB\_BLOCKING} and {\sf GrB\_NONBLOCKING} modes, the method exits with return value 
{\sf GrB\_SUCCESS} and the new contents of the output scalar is as defined above.


%-----------------------------------------------------------------------------

\subsubsection{{\sf reduce}: Matrix-scalar variant}
\label{Sec:Reduce_matrix_scalar}

Reduce all stored values into a single scalar.

\paragraph{\syntax}

\begin{verbatim}
        // scalar value + monoid (only)
        GrB_Info GrB_reduce(<type>               *val,
                            const GrB_BinaryOp    accum,
                            const GrB_Monoid      op,
                            const GrB_Matrix      A,
                            const GrB_Descriptor  desc);

        // GraphBLAS Scalar + monoid
        GrB_Info GrB_reduce(GrB_Scalar            s,
                            const GrB_BinaryOp    accum,
                            const GrB_Monoid      op,
                            const GrB_Matrix      A,
                            const GrB_Descriptor  desc);

        // GraphBLAS Scalar + binary operator
        GrB_Info GrB_reduce(GrB_Scalar            s,
                            const GrB_BinaryOp    accum,
                            const GrB_BinaryOp    op,
                            const GrB_Matrix      A,
                            const GrB_Descriptor  desc);
\end{verbatim}

\paragraph{Parameters}

\begin{itemize}[leftmargin=1.1in]
    \item[{\sf val} or {\sf s}] ({\sf INOUT}) Scalar to store final reduced value into. On input,
    the scalar provides a value that may be accumulated (optionally) with the result of the
    reduction operation.  On output, this scalar holds the results of the
    operation.

    \item[{\sf accum}] ({\sf IN}) An optional binary operator used for accumulating
    entries into existing  ({\sf s} or {\sf val})  value. If assignment rather than accumulation is
    desired, {\sf GrB\_NULL} should be specified.

    \item[{\sf op}]    ({\sf IN}) The monoid ($M = \langle D,\oplus,0 \rangle$) or binary operator 
    ($F_b = \langle D, D, D, \oplus \rangle$) used in the reduction operation. The $\oplus$ operator 
    must be commutative and associative; otherwise, the outcome of the operation is undefined.  
    
    \item[{\sf A}]     ({\sf IN}) The GraphBLAS matrix on which the
    reduction will be performed.

    \item[{\sf desc}] ({\sf IN}) An optional operation descriptor. If
    a \emph{default} descriptor is desired, {\sf GrB\_NULL} should be
    specified. Non-default field/value pairs are listed as follows:  \\

    \begin{tabular}{lllp{2.5in}}
        Param & Field  & Value & Description \\
        \hline
    \end{tabular}

    \emph{Note:} This argument is defined for consistency with the other GraphBLAS operations.
    There are currently no non-default field/value pairs that can be set for this operation.
\end{itemize}

\paragraph{Return Values}

\begin{itemize}[leftmargin=2.1in]
    \item[{\sf GrB\_SUCCESS}]         In blocking or non-blocking mode, 
    the operation completed successfully, and the
    output scalar ({\sf s} or {\sf val}) is ready to be used in the next method of 
    the sequence.

    \item[{\sf GrB\_PANIC}]           Unknown internal error.

    \item[{\sf GrB\_INVALID\_OBJECT}] This is returned in any execution mode 
    whenever one of the opaque GraphBLAS objects (input or output) is in an invalid 
    state caused by a previous execution error.  Call {\sf GrB\_error()} to access 
    any error messages generated by the implementation.

    \item[{\sf GrB\_OUT\_OF\_MEMORY}] Not enough memory available for the operation.

    \item[{\sf GrB\_UNINITIALIZED\_OBJECT}] One or more of the GraphBLAS objects 
    has not been initialized by a call to a respective constructor.

    \item[{\sf GrB\_NULL\_POINTER}]  {\sf val} pointer is {\sf NULL}.

    \item[{\sf GrB\_DOMAIN\_MISMATCH}]    The domains of input and output arguments are
    incompatible with the corresponding domains of the accumulation operator, 
    or reduce operator.
\end{itemize}

\paragraph{Description}

This variant of {\sf GrB\_reduce} computes the result of performing
a reduction across all of the stored elements of an input matrix storing the result into either
{\sf s} or {\sf val}.  This corresponds to (shown here for the scalar value case only):
\begin{equation*}
{\sf val} ~ = 
  \begin{cases}
     \hspace{3.7em} \bigoplus_{(i, j) \in \mathbf{ind}({\sf A})} {\sf A}(i, j) \text{,~~~ or} \\& \\
    {\sf val} ~\odot~ \left[ \bigoplus_{(i, j) \in \mathbf{ind}({\sf A})} {\sf A}(i, j)\right], \text{~if the the optional accumulator is specified.} 
  \end{cases}
\end{equation*}
where $\bigoplus = \bigodot({\sf op})$ and $\odot = \bigodot({\sf accum})$.  

Logically, this operation occurs in three steps:
\begin{enumerate}[leftmargin=0.85in]
\item[\bf Setup] The internal matrix used in the computation is formed 
and its domain is tested for compatibility.
\item[\bf Compute] The indicated computations are carried out.
\item[\bf Output] The result is written into the output scalar.
\end{enumerate}

One matrix argument is used in this {\sf GrB\_reduce} operation:
\begin{enumerate}
	\item ${\sf A} = \langle \bold{D}({\sf A}),\bold{size}({\sf A}),
    \bold{L}({\sf A}) = \{(i,j,A_{i,j}) \} \rangle$
\end{enumerate}

The output scalar, argument matrix, reduction operator and accumulation 
operator (if provided) are tested for domain compatibility as follows:
\begin{enumerate}

	\item If {\sf accum} is {\sf GrB\_NULL}, then $\bold{D}({\sf val})$ or $\bold{D}({\sf s})$ must be
    compatible with $\bold{D}({\sf op})$ from $M$  (or with $\bDin1({\sf op})$ and 
    $\bDin2({\sf op})$ from $F_b$).

	\item If {\sf accum} is not {\sf GrB\_NULL}, then $\bold{D}({\sf val})$ or $\bold{D}({\sf s})$ must be
    compatible with $\bDin1({\sf accum})$ and $\bDout({\sf accum})$ of the accumulation operator, and 
    $\bold{D}({\sf op})$ from $M$  (or $\bDout({\sf op})$ from $F_b$) must be compatible 
    with $\bDin2({\sf accum})$ of the accumulation operator.

	\item $\bold{D}({\sf A})$ must be compatible with $\bold{D}({\sf op})$ from $M$ 
    (or with $\bDin1({\sf op})$ and $\bDin2({\sf op})$ from $F_b$).
\end{enumerate}
Two domains are compatible with each other if values from one domain can be cast 
to values in the other domain as per the rules of the C language.
In particular, domains from Table~\ref{Tab:PredefinedTypes} are all compatible 
with each other. A domain from a user-defined type is only compatible with itself.
If any compatibility rule above is violated, execution of {\sf GrB\_reduce} ends and 
the domain mismatch error listed above is returned.

The number of values stored in the input, {\sf A}, is checked.  If there are no stored values in {\sf A}, then one of the following occurs depending on the output variant:
\[ 
{\sf {\bf L}(s)} =
\begin{cases}
    \{\}, & \mbox{(cleared) if {\sf accum} = {\sf GrB\_NULL}}, ~~\\ & \\
    {\sf {\bf L}(s)}, ~~~ ~~~~~~~ & \mbox{(unchanged) otherwise,}
\end{cases}
\]
or
\[ 
{\sf val} =
\begin{cases}
    \bold{0}({\sf op}), & \mbox{(cleared) if {\sf accum} = {\sf GrB\_NULL}}, \\ & \\
    {\sf val} ~\odot~ {\bf 0}({\sf op}), & \mbox{otherwise,}
\end{cases}
\]
where $\bold{0}({\sf op})$ is the identity of the monoid. The operation returns immediately with {\sf GrB\_SUCCESS}.

For all other cases, the internal matrix and scalar used in 
the computation is formed ($\leftarrow$ denotes copy):
\begin{enumerate}
	\item Matrix $\matrix{\widetilde{A}} \leftarrow {\sf A}$.
    \item Scalar $\scalar{\tilde{s}} \leftarrow {\sf s}$ (GraphBLAS scalar case).
\end{enumerate}
% Note that these copies are not really necessary as aliasing cannot really happen 
% but this section is here to be consistent with the format of other operations.

We are now ready to carry out the reduce and any additional associated operations.  
An intermediate scalar result $t$ is computed as follows:
\[ 
t ~ = ~
    \bigoplus\limits_{(i, j) \in \bold{ind}(\matrix{\widetilde{A}})} \matrix{\widetilde{A}}(i, j),
\]
where $\oplus = \bigodot({\sf op})$.

The final reduction value is computed as follows:
\[
{\sf {\bf L}(s)} \leftarrow
    \begin{cases}
    \{ t \}, & \text{ when {\sf accum = GrB\_NULL} or $\scalar{\tilde{s}}$ is empty, or} \\ & \\
    \{ {\bf val}(\scalar{\tilde{s}}) ~\odot ~ t \}, & \text{otherwise;} \\
    \end{cases}
\]
or 
\[
{\sf val} \leftarrow
    \begin{cases}
    t, & \text{ when {\sf accum = GrB\_NULL}, or} \hspace*{0.95in} \\ & \\
    {\sf val} ~\odot ~t, ~~~~~ & \text{otherwise;} \\
    \end{cases}
\]

In both {\sf GrB\_BLOCKING} and {\sf GrB\_NONBLOCKING} modes, the method exits with return value 
{\sf GrB\_SUCCESS} and the new contents of the output scalar is as defined above.



%=========================================================================

\subsection{{\sf transpose}: Transpose rows and columns of a matrix}

This version computes a new matrix that is the transpose of the source matrix.

\paragraph{\syntax}

\begin{verbatim}
        GrB_Info GrB_transpose(GrB_Matrix             C,
                               const GrB_Matrix       Mask,
                               const GrB_BinaryOp     accum,
                               const GrB_Matrix       A,
                               const GrB_Descriptor   desc);
\end{verbatim}

\paragraph{Parameters}

\begin{itemize}[leftmargin=1.1in]
    \item[{\sf C}]    ({\sf INOUT}) An existing GraphBLAS matrix. On input,
    the matrix provides values that may be accumulated with the result of the
    transpose operation.  On output, the matrix holds the results of the
    operation.

    \item[{\sf Mask}] ({\sf IN}) An optional ``write'' mask that controls which
    results from this operation are stored into the output matrix {\sf C}. The 
    mask dimensions must match those of the matrix {\sf C}. If the 
    {\sf GrB\_STRUCTURE} descriptor is {\em not} set for the mask, the domain of the 
    {\sf Mask} matrix must be of type {\sf bool} or any of the predefined 
    ``built-in'' types in Table~\ref{Tab:PredefinedTypes}.  If the default
    mask is desired (\ie, a mask that is all {\sf true} with the dimensions of {\sf C}), 
    {\sf GrB\_NULL} should be specified.

    \item[{\sf accum}] ({\sf IN}) An optional binary operator used for accumulating
    entries into existing {\sf C} entries.
    If assignment rather than accumulation is
    desired, {\sf GrB\_NULL} should be specified.

    \item[{\sf A}]     ({\sf IN}) The GraphBLAS matrix on which
    transposition will be performed.

    \item[{\sf desc}] ({\sf IN}) An optional operation descriptor. If
    a \emph{default} descriptor is desired, {\sf GrB\_NULL} should be
    specified. Non-default field/value pairs are listed as follows:  \\

    \hspace*{-2em}\begin{tabular}{lllp{2.7in}}
        Param & Field  & Value & Description \\
        \hline
        {\sf C}    & {\sf GrB\_OUTP} & {\sf GrB\_REPLACE} & Output matrix {\sf C}
        is cleared (all elements removed) before the result is stored in it.\\

        {\sf Mask} & {\sf GrB\_MASK} & {\sf GrB\_STRUCTURE}   & The write mask is
        constructed from the structure (pattern of stored values) of the input
        {\sf Mask} matrix. The stored values are not examined.\\

        {\sf Mask} & {\sf GrB\_MASK} & {\sf GrB\_COMP}   & Use the 
        complement of {\sf Mask}. \\

        {\sf A}    & {\sf GrB\_INP0} & {\sf GrB\_TRAN}   & Use transpose of {\sf A}
        for the operation. \\
    \end{tabular}
\end{itemize}

\paragraph{Return Values}

\begin{itemize}[leftmargin=2.1in]
    \item[{\sf GrB\_SUCCESS}]         In blocking mode, the operation completed
    successfully. In non-blocking mode, this indicates that the compatibility 
    tests on dimensions and domains for the input arguments passed successfully. 
    Either way, output matrix {\sf C} is ready to be used in the next method of
    the sequence.

    \item[{\sf GrB\_PANIC}]           Unknown internal error.

    \item[{\sf GrB\_INVALID\_OBJECT}] This is returned in any execution mode 
    whenever one of the opaque GraphBLAS objects (input or output) is in an invalid 
    state caused by a previous execution error.  Call {\sf GrB\_error()} to access 
    any error messages generated by the implementation.

    \item[{\sf GrB\_OUT\_OF\_MEMORY}] Not enough memory available for the operation.

    \item[{\sf GrB\_UNINITIALIZED\_OBJECT}] One or more of the GraphBLAS objects 
    has not been initialized by a call to {\sf new} (or {\sf Matrix\_dup} for matrix
    parameters).

    \item[{\sf GrB\_DIMENSION\_MISMATCH}]  {\sf mask}, {\sf C} and/or {\sf A} 
    dimensions are incompatible.

    \item[{\sf GrB\_DOMAIN\_MISMATCH}]    The domains of the various matrices are
    incompatible with the corresponding domains of the 
    accumulation operator, or the mask's domain is not compatible with {\sf bool}
    (in the case where {\sf desc[GrB\_MASK].GrB\_STRUCTURE} is not set).
\end{itemize}

\paragraph{Description}

{\sf GrB\_transpose} computes the result of performing a transpose of the input matrix:
${\sf C} = {\sf A}^T$; or, if an optional binary accumulation 
operator ($\odot$) is provided, ${\sf C} = {\sf C} \odot {\sf A}^T$.  
We note that the input matrix {\sf A} can itself be optionally transposed before the operation,
which would cause either an assignment from {\sf A} to {\sf C} or an
accumulation of {\sf A} into {\sf C}.

Logically, this operation occurs in three steps:
\begin{enumerate}[leftmargin=0.75in]
\item[\bf Setup] The internal matrix and mask used in the computation are formed 
and their domains and dimensions are tested for compatibility.
\item[\bf Compute] The indicated computations are carried out.
\item[\bf Output] The result is written into the output matrix, possibly under 
control of a mask.
\end{enumerate}

Up to three matrix arguments are used in this {\sf GrB\_transpose} operation:
\begin{enumerate}
	\item ${\sf C} = \langle \bold{D}({\sf C}),\bold{nrows}({\sf C}),
    \bold{ncols}({\sf C}),\bold{L}({\sf C}) = \{(i,j,C_{ij}) \} \rangle$

	\item ${\sf Mask} = \langle \bold{D}({\sf Mask}),\bold{nrows}({\sf Mask}),
    \bold{ncols}({\sf Mask}),\bold{L}({\sf Mask}) = \{(i,j,M_{ij}) \} \rangle$ (optional)

	\item ${\sf A} = \langle \bold{D}({\sf A}),\bold{nrows}({\sf A}),
    \bold{ncols}({\sf A}),\bold{L}({\sf A}) = \{(i,j,A_{ij}) \} \rangle$
\end{enumerate}

The argument matrices and accumulation 
operator (if provided) are tested for domain compatibility as follows:
\begin{enumerate}
	\item If {\sf Mask} is not {\sf GrB\_NULL}, and ${\sf desc[GrB\_MASK].GrB\_STRUCTURE}$
    is not set, then $\bold{D}({\sf Mask})$ must be from one of the pre-defined types of 
    Table~\ref{Tab:PredefinedTypes}.

	\item $\bold{D}({\sf C})$ must be
    compatible with $\bold{D}({\sf A})$ of the input matrix.

	\item If {\sf accum} is not {\sf GrB\_NULL}, then $\bold{D}({\sf C})$ must be
    compatible with $\bDin1({\sf accum})$ and $\bDout({\sf accum})$ of the accumulation operator and 
    $\bold{D}({\sf A})$ of the input matrix must be compatible with $\bDin2({\sf accum})$ of the accumulation operator.

\end{enumerate}
Two domains are compatible with each other if values from one domain can be cast 
to values in the other domain as per the rules of the C language.
In particular, domains from Table~\ref{Tab:PredefinedTypes} are all compatible 
with each other. A domain from a user-defined type is only compatible with itself.
If any compatibility rule above is violated, execution of {\sf GrB\_transpose} ends and 
the domain mismatch error listed above is returned.

From the argument matrices, the internal matrices and mask used in 
the computation are formed ($\leftarrow$ denotes copy):
\begin{enumerate}
	\item Matrix $\matrix{\widetilde{C}} \leftarrow {\sf C}$.

	\item Two-dimensional mask, $\matrix{\widetilde{M}}$, is computed from
    argument {\sf Mask} as follows:
	\begin{enumerate}
		\item If ${\sf Mask} = {\sf GrB\_NULL}$, then $\matrix{\widetilde{M}} = 
        \langle \bold{nrows}({\sf C}), \bold{ncols}({\sf C}), \{(i,j), 
        \forall i,j : 0 \leq i <  \bold{nrows}({\sf C}), 0 \leq j < 
        \bold{ncols}({\sf C}) \} \rangle$.

		\item If {\sf Mask} $\ne$ {\sf GrB\_NULL},
        \begin{enumerate}
            \item If ${\sf desc[GrB\_MASK].GrB\_STRUCTURE}$ is set, then 
            $\matrix{\widetilde{M}} = \langle \bold{nrows}({\sf Mask}), 
            \bold{ncols}({\sf Mask}), \{(i,j) : (i,j) \in \bold{ind}({\sf Mask}) \} \rangle$,
            \item Otherwise, $\matrix{\widetilde{M}} = \langle \bold{nrows}({\sf Mask}), 
            \bold{ncols}({\sf Mask}), \\ \{(i,j) : (i,j) \in \bold{ind}({\sf Mask}) \wedge 
            ({\sf bool}){\sf Mask}(i,j) = \true\} \rangle$.
        \end{enumerate}

		\item	If ${\sf desc[GrB\_MASK].GrB\_COMP}$ is set, then 
        $\matrix{\widetilde{M}} \leftarrow \neg \matrix{\widetilde{M}}$.
	\end{enumerate}

	\item Matrix $\matrix{\widetilde{A}} \leftarrow
    {\sf desc[GrB\_INP0].GrB\_TRAN} \ ? \ {\sf A}^T : {\sf A}$.
\end{enumerate}

The internal matrices and masks are checked for dimension compatibility. The following
conditions must hold:
\begin{enumerate}
	\item $\bold{nrows}(\matrix{\widetilde{C}}) = \bold{nrows}(\matrix{\widetilde{M}})$.

	\item $\bold{ncols}(\matrix{\widetilde{C}}) = \bold{ncols}(\matrix{\widetilde{M}})$.

	\item $\bold{nrows}(\matrix{\widetilde{C}}) = \bold{ncols}(\matrix{\widetilde{A}})$.

	\item $\bold{ncols}(\matrix{\widetilde{C}}) = \bold{nrows}(\matrix{\widetilde{A}})$.
\end{enumerate}
If any compatibility rule above is violated, execution of {\sf GrB\_transpose} ends and
the dimension mismatch error listed above is returned.

From this point forward, in {\sf GrB\_NONBLOCKING} mode, the method can 
optionally exit with {\sf GrB\_SUCCESS} return code and defer any computation 
and/or execution error codes.

We are now ready to carry out the matrix transposition and any additional 
associated operations.  We describe this in terms of two intermediate matrices:
\begin{itemize}
    \item $\matrix{\widetilde{T}}$: The matrix holding the  transpose of
    $\matrix{\widetilde{A}}$.
    \item $\matrix{\widetilde{Z}}$: The matrix holding the result after 
    application of the (optional) accumulation operator.
\end{itemize}

The intermediate matrix
\[
\matrix{\widetilde{T}} = \langle
\bold{D}(\matrix{\sf A}), \bold{ncols}(\matrix{\widetilde{A}}),
\bold{nrows}(\matrix{\widetilde{A}}), 
\{(j,i,A_{ij}) \forall (i,j) \in \bold{ind}(\matrix{\widetilde{A}}) 
\rangle
\]
is created.  

\input{ops_accum_z_matrix}

\input{ops_mask_replace_matrix}
