\section{Standard Definitions}
\label{Sec:stddef}

\newcommand{\integer}{{\sf int32\_t}}
\newcommand{\float}{{\sf float}}
\newcommand{\double}{{\sf double}}
\newcommand{\bool}{{\sf bool}}
\newcommand{\false}{{\sf false}}
\newcommand{\true}{{\sf true}}

The current GraphBLAS specification could be augmented with a list of 
pre-defined objects, such as descriptors, semirings and monoids, that are commonly used
in building graph algorithms. These standard pre-defined objects would
simplify coding and ensure more consistency across algorithms. 
We emphasize
that individual application writers are always free to create their own set of
pre-defined objects.

Additionally, GraphBLAS could include pre-defined macros that support
shorter calling syntax for methods, by hiding default arguments. 
Again, individual application writers are always free to create their
own set of pre-defined objects.

We propose that such pre-defined objects and macros would
be included in a compilation unit through a construct like:

\begin{verbatim}
#include "GrB_stddef.h"
\end{verbatim}

Table~\ref{Tab:stddef} shows an initial set of proposed standard object
definitions.  It includes descriptors for transposing either or both input
matrices, booleam monoids and the standard arithmetic monoids for integer,
single-precision floating-point and double-precision floating-point
data.  It also includes the conventional boolean semiring, as well as
arithmetic semirings for integer, single-precision floating-point and
double-precision floating-point data.  Other semirings commonly used
with graph algorithms can also be added.

\begin{table}[htb]
	\hrule
	\caption{Initial set of proposed standard definitions for GraphBLAS.}
	\label{Tab:stddef}
	\begin{center}
	\begin{tabular}{|l|l|l|} \hline
		type		& name				& description \\ \hline
				& {\sf GrB\_TA}			& ${\sf GrB\_INP0} = {\sf GrB\_TRAN}$ \\
		Descriptor	& {\sf GrB\_TB}			& ${\sf GrB\_INP1} = {\sf GrB\_TRAN}$ \\ 
				& {\sf GrB\_TAB}		& ${\sf GrB\_INP0} = {\sf GrB\_INP1} = {\sf GrB\_TRAN}$ \\ \hline
				& {\sf GrB\_Lor}		& $\left\langle \bool, \vee, \false \right\rangle$ \\
				& {\sf GrB\_Land}		& $\left\langle \bool, \wedge, \true \right\rangle$ \\
				& {\sf GrB\_Int32Add}		& $\left\langle \integer, +, 0 \right\rangle$ \\
		Monoid		& {\sf GrB\_Int32Mul}		& $\left\langle \integer, \times, 1 \right\rangle$ \\
				& {\sf GrB\_FP32Add}		& $\left\langle \float, +, 0.0 \right\rangle$ \\
				& {\sf GrB\_FP32Mul}		& $\left\langle \float, \times, 1.0 \right\rangle$ \\
				& {\sf GrB\_FP64Add}		& $\left\langle \double, +, 0.0 \right\rangle$ \\ 
				& {\sf GrB\_FP64Mul}		& $\left\langle \double, \times, 1.0 \right\rangle$ \\ \hline
				& {\sf GrB\_Boolean}		& $\left\langle \bool, \bool, \bool, \vee, \wedge, \false \right\rangle$ \\
		Semiring	& {\sf GrB\_Int32AddMul}	& $\left\langle \integer, \integer, \integer, +, \times, 0 \right\rangle$ \\
				& {\sf GrB\_FP32AddMul}		& $\left\langle \float, \float, \float, +, \times, 0.0 \right\rangle$ \\ 
				& {\sf GrB\_FP64AddMul}		& $\left\langle \double, \double, \double, +, \times, 0.0 \right\rangle$ \\ \hline
	\end{tabular}
	\end{center}
	\hrule
\end{table}
